### 2023-03-28 alloscore_talk/pres.qmd
getting started

---
title: "alloscore talk"
format:
  revealjs: 
    width: 1500
editor: visual
---

## Motivation

Gov to allocate scarce hospital supplies across 50 states

-   given many (probabilistic) forecasts of need, which should they use?

Raises question: how will chosen forecast be used?

One possibility: to minimize $E_F \sum_{l=1}^{51} \max(0, Y_i - x_i)$

-   $Y_{1:51} \sim F$ are future needs ($F$ unknown)
-   $x_{1:51}$ are allocated quantities
-   only $K$ units of supplies are available

## A constrained stochastic optimization problem

Let $\overline{s}_F(\mathbf{x}) = E_F \sum \max(0, Y_i - x_i)$

Lagrange multiplier equation: $\nabla \overline{s}_F(\mathbf{x}) + \lambda =0$

$$
\frac{d}{dx_i}E_F \max(0, Y_i - x_i) = \frac{d}{dx_i}\int_{x_i}^{\infty} (y - x_i)f_i(y)dy = F_i(x_i)-1 = -\lambda\\
x_i = F_i^{-1}(1-\lambda)
$$

Solve $\sum_{l=1}^{51} F_i^{-1}(1-\lambda) = K$ for $\lambda^{\star} \in (0,1)$ to get best allocation $x_i^{\star} = F_i^{-1}(1-\lambda^{\star})$

## Solution for location-scale family $F_i$'s

$$
\sum F_i^{-1}(1-\lambda) = \sum \mu_i + \sigma_i \Phi^{-1}(1-\lambda )= K \\
\implies \lambda^{\star} = 1 - \Phi\left(\frac{K-\sum \mu_j}{\sum \sigma_j}\right) \\
\implies x_i^{\star} = 
F_i^{-1}\circ\Phi\left(\frac{K-\sum \mu_j}{\sum \sigma_j}\right)
=\mu_i+\sigma_i\left(\frac{K-\sum \mu_j}{\sum \sigma_j}\right)
$$

Interpretation: underdispersion at loc $i$ -\> too little of "excess or shortage in mean" $K-\sum \mu_j$ given to or removed from $i$

## Allocation as forecast perfomance

$\mathbf{x}_F^{\star}$ is a **Bayes act** for $F$ under loss $$
s(\mathbf{x}, \mathbf{y}) = \sum \max(0,y_i - x_i)
$$ The quality of forecast $F$ then can be estimated by how much loss using it actually incurs, i.e., $$
S(\mathbf{F}, \mathbf{y}) = \sum \max(0,y_i - x_{F,i}^{\star})
$$

-   or perhaps $S(\mathbf{F}, \mathbf{y}) - \max(0, \sum y_i-K)$, performance against an oracle

## Scoring function for LS case

$$
s_K(\mathbf{Q}, \mathbf{y})=\sum \mathbb{1}\left\{\frac{K-\sum \mu_i}{\sum \sigma_i} \leq \frac{y_i-\mu_i}{\sigma_i}\right\}\left(y_i-x_{F,i}^{\star}\right)
$$

## Scoring Functions in general

Forecaster $A$ asked to estimate a **functional** $T(F_Y)$ for a RV $Y\sim F_Y$.

-   mean, median, variance, etc.
-   $A$ believes that $Y \sim F$

After $A$ gives us $x$ and nature realizes $Y=y$, we judge performance with a **scoring function** $s(x,y)$

$s$ is **(strictly) consistent** for $T$ if$A$ cannot expect to do better than offer $x=T(F)$ as a forecast of $T(F_Y)$, i.e,

$$
E_F[s(T(F)),Y] (<) \leq E_F[s(x,Y)] \text{ for all } F,x
$$

## Elicitability

A functional is **elicitable** if it has a strictly consistent scoring function.

Basic example are $\alpha$ quantiles, with SF:

$$
\begin{aligned}
s(x, y)=s_{O, U}(x, y) &=O(g(x)-g(y))_{+}+U(g(x)-g(y))_{-} . \\
&=\kappa\left((1-\alpha)(g(x)-g(y))_{+}+\alpha(g(x)-g(y))_{-}\right) \\
&=\kappa(\mathbb{1}\{x>y\}-\alpha)(g(x)-g(y)) \\
&:=s_{\kappa, \alpha}(x, y):=\kappa s_\alpha(x, y)
\end{aligned}
$$

$\alpha=U /(U+O), \kappa = O+U$

## Reformulation of allocation problem

Hosp allocation problem of this form in each location, with $O_i=0, U_i=1$

With constraint $x \leq K$ Bayes act is just $Q=\min(K,F^{-1} (U/(U+O)))=K$

But taken jointly, have non-trivial BA $Q_i = F_i^{-1}\left(\alpha_i\left(1-\lambda_K / U_i\right)\right)$ or(!) 0

-   $\lambda_K =0$ corresponds to case when $\sum Q_i(\alpha_i) \leq K$

## Domain versions

### Newsvendor problem

Ordering decision faced by a retailer of a perishable good (such as newspapers) when customer demand is uncertain.

-   Allocation problem appears as **multi-product NVP with stocking constraint**
-   Well-studied in OR literature

### Meteoroligist/Epidemiologist Cost/Loss Problem

-   have not seen much in allocation form

No literature regarding allocation solution as scoring rule

## Solving Allocation Problem

Straightforward in simple situations

-   all $F_i$ smooth with same support starting at 0
-   $K$ not too much smaller than $\mathbf{w}^T \mathbf{Q}(\alpha)$

$$
\nabla Z(\mathbf{x}) + \lambda \mathbf{w} =0 \\
\implies \kappa_i(F_i(Q_i) - \alpha) + \lambda w_i = 0 \\
\implies Q_i(\lambda) = F_i^{-1}(\alpha_i(1 - \lambda w_i / U_i))
$$

Now solve $\sum_{i=1}^N w_i Q_i(\lambda)=K$ for $\lambda^{\star}$ to get optimal allocation $\mathbf{Q}(\lambda^{\star})$.

## Complications

If some $Q_i$ need to be 0, we don't know in general which set of 0's is best

-   problem becomes intractable for large $N$
-   need a non-linear programming algorithm... `alloscore` implements one

## Scoring Rules

-   Should be regret-based, i.e., how well you do against an oracle that knows $y$'s and never wastes any resources

## Next Steps

-   Use Evan's package [`distfromq`](https://github.com/reichlab/distfromq) to interpolate quantile functions and distribtuions from hub quantile forecasts and score across $K$ range

-   extend this to no overprediction penalty i.e. $\alpha_i = 1$ scenario

## Aspirations

-   Allocation skill could be an indicator for peak/outbreak detection skill