### 2023-04-18 notes/crps_connection_exp.Rmd

---
title: "Connection between allocation scoring rules an CRPS: example with exponential forecasts"
date: "`r Sys.Date()`"
output: 
    pdf_document:
        latex_engine: xelatex
urlcolor: blue
bibliography: allocation_notes.bib 
header-includes:
   - \usepackage{amsmath}
   - \usepackage{algorithm}
   - \usepackage[noend]{algpseudocode}
   - \usepackage{amsthm}
   - \usepackage{bm}
   - \usepackage{cases}
   - \usepackage{caption}
   - \usepackage{unicode-math}
   - \usepackage{hyperref}
   - \DeclareMathOperator*{\argmin}{argmin}
---

\newcommand{\sbar}{\overline{s}}

```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
```

Consider a simplified setting where $O_i = 0$ and $U_i = U$ are shared for all $i$, and $g_i(x_i) = x_i$. This is the setting we have laid out in the initial alloscore manuscript. We consider three settings:

 1. The decision maker is fixed with a fixed, known, constraint on the total allocation, $K$.
 2. The decision maker has some uncertainty about the total constraint $K$.
 3. We would like to understand the relationship between the allocation score and CRPS by exhibiting the distribution on $K$ that would lead to an "equally-weighted" CRPS.

Throughout, we will suppose there are $n = 2$ locations, and a forecaster produces the forecasts $Y_1 \sim Exp(1/\sigma_1)$ with $\sigma_1 = 1$ and $Y_2 \sim Exp(1 / \sigma_2)$ with $\sigma_2 = 5$. The quantile functions corresponding to these forecasts are given by $F_i^{-1}(\tau) = -\sigma_i \log(1 - \tau)$, where $\tau$ is a probability level which we take to be in $(0, 1)$.

The expected loss function is
\begin{equation*}
\bar{s}_F(\mathbf{x}) = \sum_{i=1}^2 \mathbb{1}\{x_i < 0\}(\sigma - x_i) + \mathbb{1}\{0 \leq x_i\}\sigma e^{-x_i/\sigma}.
\end{equation*}

\section{Fixed $K$}

At a fixed constraint $K$, in our simplified setting the solution of the allocation problem is given by the quantiles $(F_1^{-1}(\tau), F_2^{-1}(\tau))$ at a probability level $\tau$ (connecting to notation elsewhere, $\tau = 1 - \lambda^\star$) such that
\begin{align*}
K &= F_1^{-1}(\tau) + F_2^{-1}(\tau) \\
  &= -\log(1 - \tau) (\sigma_1 + \sigma_2) \\
  &= h(\tau)
\end{align*}
We can rearrange to obtain
\begin{equation}
\tau = h^{-1}(K) = 1 - \exp[-K/(\sigma_1 + \sigma_2)]. \label{eqn:tau_fn_K}
\end{equation}

We can visualize this in terms of the expected loss function (shown in shades of blue and yellow) as follows:

```{r, cache = TRUE}
sigma_1 <- 1
sigma_2 <- 5

K <- 5

tau <- 1 - exp(-K/(sigma_1 + sigma_2))

x_1star <- qexp(tau, rate = 1 / sigma_1)
x_2star <- qexp(tau, rate = 1 / sigma_2)

n_grid <- 101
grid_l <- -2.0
grid_u <- 9.0
y_grid <- tidyr::expand_grid(
    y_1 = seq(from = grid_l, to = grid_u, length.out = n_grid),
    y_2 = seq(from = grid_l, to = grid_u, length.out = n_grid)
)

s_bar <- function(x_1, x_2) {
    sigma_1 * exp(-x_1 / sigma_1)*(x_1>=0) + sigma_2 * exp(-x_2 / sigma_2)*(x_2>=0) +
    (sigma_1 - x_1)*(x_1<0) + (sigma_2 - x_2)*(x_2<0) 
}

joint_dist <- y_grid %>%
    dplyr::mutate(
        s_bar = s_bar(y_1, y_2)
    )

ggplot(data = joint_dist) +
    geom_raster(aes(x = y_1, y = y_2, fill = s_bar)) +
    geom_contour(mapping = aes(x = y_1, y = y_2, z = s_bar), breaks = s_bar(x_1star, x_2star)%%1 + seq(0, 10, by = .5)) +
    geom_abline(intercept = K, slope = -1) +
    geom_point(x = x_1star, y = x_2star, color = "orange") +
    scale_fill_viridis_c() +
    xlim(grid_l, grid_u) +
    ylim(grid_l, grid_u) +
    theme_bw()
```

\section{Obtaining the CRPS weighting associated with a distribution on $K$}

Suppose that the decision maker has some uncertainty about the level of the constraint $K$, expressed by the distribution $F_K$.  For concreteness, here we take this distribution to be $Gamma(500, 0.01)$ using a shape and scale parameterization. Here's a picture of this distribution, which is concentrated near $K = 5$:

```{r}
curve(dgamma(x, shape = 500, scale = 0.01), from = 0, to = 10, n = 1001)
```

As discussed above, given fixed forecasts $F_1$ and $F_2$, for each value of the constraint $K$, a quantile probability level $\tau = h^{-1}(K) = 1 - \exp[-K/(\sigma_1 + \sigma_2)]$ is determined. We can therefore use a change of variables to obtain a density for $\tau$ from the density for $K$ as follows:

\begin{align*}
f_T(\tau) &= f_{K}(h(\tau)) \left\vert \frac{d}{d\tau} h(\tau) \right\vert \\
&= f_{K}(-\log(1 - \tau) (\sigma_1 + \sigma_2)) \frac{(\sigma_1 + \sigma_2)}{1 - \tau} \\
\end{align*}

Here's a plot of this induced density on $\tau$:

```{r}
tau <- seq(from = 0.001, to = 0.999, length.out = 10000)

d_tau <- function(tau) {
    dgamma(-log(1 - tau) * (sigma_1 + sigma_2), shape = 500, scale = 0.01) *
                ((sigma_1 + sigma_2) / (1 - tau))
}

plot(tau, d_tau(tau), "l")

# note that this is a density
sum(d_tau(tau) * diff(tau)[1])
```

We can think of the allocation score determined by $F_K$ as corresponding to a weighted CRPS with the weighting expressed by the above density on quantile levels. However, note that this interpretation is specific to this forecast. A different forecast would translate to a different weighting on quantile levels.

\section{Reproducing equally weighted CRPS}

Going in the other direction, given forecasts $F_1$ and $F_2$, we can determine the distribution on values of the constraint $K$ that corresponds to an equal weighting of all quantile levels, as is done in CRPS.  Again, this implied distribution on $K$ depends on the forecasts and will be different for different forecasters.

In this setting, we start with $\tau \sim Unif(0, 1)$, and note that given a value of $\tau$ we can calculate $K = h(\tau) = -\log(1 - \tau) (\sigma_1 + \sigma_2)$. This induces a distribution on $K$ with density
\begin{align*}
f_K(K) &= f_{T}(h^{-1}(K)) \left\vert \frac{d}{dK} h^{-1}(K) \right\vert \\
&= 1 \cdot (\sigma_1 + \sigma_2)^{-1} \exp[-K/(\sigma_1 + \sigma_2)]
\end{align*}
This is the density of an $Exp(\sigma_1 + \sigma_2)$ random variable.

Thus, with our forecasts, CRPS corresponds to the weighting on resource constraints illustrated in the following figure:

```{r}
curve(dexp(x, rate = 1 / (sigma_1 + sigma_2)), 0, 20)
```

This corresponds to placing large mass on small values of the constraint $K$, which may not correspond well to actual knowledge about the resource constraints.