### 2023-06-12 ISF_talk/talk.qmd
add Norm1, a normal simulation experiment

---
title: "alloscore, ISF 43"
format:
  revealjs: 
    width: 1500
editor: visual
---

Comparitive vs absolute assessment/evaluation

illustrate algo with a lambda over xs plot against tau