### 2023-06-14 ISF_talk/talk.qmd
more entropy/divergence work.  getting slide started.

---
title: "alloscore, ISF 43"
format:
  revealjs: 
    width: 1500
editor: visual
---

## The backstory

Over last decade, multicenter effort to collect forecasts via hubs for ID outbreaks (epi/pan/endemic)
with central purpose to "inform public health responses"

With uncertainty quantification being well-established as needed DM input, 
hubs have trended toward eliciting probabilistic forecasts

Basic element of any such response is an *allocation decision* regarding
 - medical supplies
 - personnel 
 - facility capacity
 - R&D funding
 - public tolerance for NPIs

Allocations can be wrt space, time, demographic

Similar to other contexts where there are large-scale institutional efforts to use expert judgment to assess and manage social risk
 - finance
 - meteorology
 - military conflict

Even with complex game-theoretic approachs to collective decision making,
*allocation* almost always a requirement to strategy development

## Scoring rules and social utility

Hubs have tried to use SRs to rank and combine forecasts in ways that offer best suppport for PH responders

This can be cast in decision-theoretic framework of utility, loss functions, and Bayes acts.

The DM needs to decide what action $a$ to take regarding an uncertain outcome $Y$ such as the number of severe cases in various locations.
Suppose the DM also has
a loss function $l(a,y)$, e.g., the unmet need when $a$ beds are allocated for $y$ severe cases,
and a utility function U quantifying, e.g., the ratio of the value of preventing 1000 unhospitalized severe cases versus the value of preventing 10 unhospitalized severe cases.  That is U specifies numbers such as U(l(10,20)) / U(l(1000,2000)).
Then the Bayes act relative to a forecast $Y \sim F$ is the action $a^F$ that will maximize the DM’s expected utility assuming that in reality $Y \sim F$.

## Basic computational approach

Adopt an iterative Lagrange multiplier method following long history in inventory management of solving
*constrained multi-product newsvendor* problems

illustrate algo with a lambda over xs plot against tau

Comparitive vs absolute assessment/evaluation