### 2023-06-16 ISF_talk/talk.qmd
current state of isf talk

---
title: "alloscore, ISF 43"
format:
  revealjs: 
    width: 1500
editor: visual
---

## Timeline

::: notes
West Nile mosq control decisions
:::

## The Backstory

Starting \~ 2010? Outbreak forecasting hubs developed to "inform public health responses"

Trend toward elicitation of probabilistic forecasts

 - uncertainty quantification often essential to a public health **DM** (decision maker)

But in the inimitable words of chatGPT:

> It's important to note that while these efforts represent significant advances in the field, forecasting the spread and impact of infectious diseases is a very complex task with many challenges, and these models should be interpreted with caution.

::: notes
Over last two decades, multicenter effort to collect forecasts via hubs for ID outbreaks (epi/pan/endemic) with central purpose to "inform public health responses"

With uncertainty quantification being well-established as needed DM input, 
:::

## The fundamental need to decide allocations

::: columns
::: {.column width="40%"}
Basic element of any such response is an *allocation decision* regarding

 - medical supplies 
 - personnel 
 - facility capacity 
 - R&D funding 
 - public tolerance for NPIs
:::

::: {.column width="60%"}
Similar to other contexts where there are large-scale institutional efforts to use expert judgment to assess and manage social risk

 - finance 
 - meteorology 
 - military conflict
:::
:::

::: notes
Allocations can be wrt space, time, demographic

Even with complex game-theoretic approachs to collective decision making, *allocation* almost always a requirement to strategy development
:::

## Scoring rules and social utility

Hubs have tried to use SRs to rank and combine forecasts in ways that offer best suppport for PH responders

This can be cast in decision-theoretic framework of utility, loss functions, and Bayes acts.

The DM needs to decide take an action $a$ wrt an outcome $Y$

- such as the number of severe cases in various locations.

Suppose the DM also has a *loss function* $l(a,y)$ - e.g., the lost social/political/economic/psychological/moral utility caused by unmet need when $a < y$ beds are allocated and $y$ severe cases occur - this must somehow acknowledge risk attitude heterogeneity among various DM's that a forecast hub might serve

Then the *Bayes act* relative to a forecast $Y \sim F$ is the action $a^F$ that will minimize the DM's expected loss assuming that in reality $Y \sim F$.

------------------------------------------------------------------------

A Scoring Rule attempts to transfer or "share" this loss function with the forecaster through a numerical reward system where FP receives

$$
S(F,y)
$$

units of money, goodwill, prestige, etc. A central goal in designing $S$ is to create a scenario where the FP beleives she cam maximizes her own expected utility by giving the DM the forecast that the FP believes in actually true and/or best suited for leading the DM toward actions that will maximize their utility. An SR that acheieves this is called *proper* and Bayes acts form a linchpin of their operation since the establish an explicit vacabulary of loss in place of vague efforts to rank preformance of various statisticla funcitonals adoptic form statistic inference

## Loss vs Utility

Change in loss but not utility from change in - variant severity (assuming target is infections and decision is number of beds)

Change in utility but not in loss from change in - supply and personnel cost structures - public opinion regarding importance of fatality prevention - knowledge of demographic disparities in consequences of outcomes

## Basic computational approach

Adopt an iterative Lagrange multiplier method following long history in inventory management of solving *constrained multi-product newsvendor* problems

illustrate algo with a lambda over xs plot against tau

## Scoring against an oracle

To get more interpretable allocation scores we adjust by the score of an oracle that allocates *fairly* based on what demands actually occur

Note: The oracle is not the DGP, it is a time-traveller

Comparitive vs absolute assessment/evaluation