### 2023-06-16 talks/talk2/pres2.qmd

---
title: "alloscore talk 2"
format:
  revealjs: 
    width: 1500
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Agenda

Intuitive understanding of expected unmet/overmet need

-   i.e. of "hockey-stick" or "piece-wise linear" loss $\max(0,Y - x)= (Y-x)_+$ and $\min(0,Y - x)= (Y-x)_-$

-   more broadly, quantile scores and therefore also WIS, CRPS, and who knows what else

-   a meta-theorem going back to Savage: all loss functions (for estimation or prediction) are enhancements of either the squared-error or piece-wise linear error.

My own experience: the more I tried to think about this, the more I realized I did not understand what a mean of a RV is.

## The "survival formula" for expectation

For a non-negative RV $Y$

$$
\mathbb{E}[Y]=\int_0^{\infty} \mathbb{P}(Y>y) \mathrm{d} y = \int_0^{\infty}(1- F_Y(y))dy
$$

Appears (unhelpfully) as an exercise at the beginning of many math-stats books

## Basic idea

Take sample points $y_i = ih$ and assume $Nh$ large enough that $F_Y(Nh)$ close to 1. 
$$
\mathbb{E}[Y] \approx \sum_{i=1}^{N} y_if_Y(y_i) = \sum_{i=1}^{N}ih f_Y(ih) = \\
\begin{array} 
& & & & & & \vdots \\
& & & & hf_Y(3h) + \cdots\\
& & hf_Y(2h) &+ &hf_Y(3h) + \cdots\\
 hf_Y(h) &+ &hf_Y(2h) &+ &hf_Y(3h) + \cdots
\end{array}\\
\ \\
\approx \int_0^{\infty}(1- F_Y(y))dy
$$

## Resclicing picture

```{r echo=FALSE}
f <- function(y) 5*(.3*dnorm(y-3, sd = .8) + .7*dnorm(y-7,sd=2))
fplt <- function(y) 10*f(y)
x0 <- -2
xl <- 4
xu <- 15
width <- xu-xl
N <- 20
h <- width/N

# dat <- tibble(yis = seq(x0,xu,by=h), heights = pmax(0,yis-xl), fis = f(yis)*(heights>0), fyis = heights*fis)
dat <- tibble(
  yis = seq(xl+h,xu,by=h), 
  heights = yis-xl, 
  fis = f(yis), 
  fyis = heights*fis
  )
y1 <- dat$yis[1]
ylast <- tail(dat$yis,1)

p <- ggplot(data = dat) + geom_function(fun = fplt, n = 1000) +
  geom_function(fun = function(y) pmax(0,y-xl)) +
  scale_y_continuous(breaks = dat$heights) +
  scale_x_continuous(breaks = dat$yis, limits = c(x0,xu)) +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        panel.grid.major.y = element_line(color = "lightgrey")
        )
p

p <- p + geom_rect(aes(xmin=yis-.1*h, xmax=yis+.1*h, ymin=0, ymax=heights), alpha=.2)
p
p <- p + geom_rect(aes(xmin=yis+.15*h, xmax=yis+.35*h, ymin=0, ymax=fyis)) 
p
p <- p + map(1:N, function(idx){
      geom_rect(data = dat %>% slice_tail(n = N + 1 - idx), 
                      aes(xmin=.data$yis+.4*h, 
                          xmax=.data$yis+.6*h, 
                          ymin=.data$heights[1] - h, 
                          ymax=.data$heights[1] - h + h*.data$fis))
    })
p

# p <- p + map(1:N, ~geom_line(
#     data = tibble(x=seq(xl+h*., ylast, length.out = 500), y = dat$yis[.]-y1+h*f(x)),
#     aes(x=x,y=y)
#     )) 
# p
```

##

So 
$$
\mathbb{E}[(Y-x)_+]= \int_x^{\infty}(1- F_Y(y))dy
$$