### 2023-06-28 talks/ISF_talk/talk.qmd
post ISF

---
title: |
  | Infectious Disease Forecast Evaluation Based On Realized Social Utility:
  | Allocation Scores
  | ISF 43
author: Aaron Gerding, Nick Reich, Ben Rogers, Evan Ray
institute: "School of Public Health and Health Sciences, UMass, Amherst"
format:
  revealjs: 
    width: 1050
    height: 700
revealjs-plugins:
  - attribution
editor: visual
css: styles.css
---

## Introduction

::: columns

::: {.column width="50%"}

About me:

- PhD in pure math
- transitioned to public health data science 
- started as RA with Nick and Evan at [reichlab.io](reichlab.io) at UMass in 2021
- maintain COVID-19 Forecast Hub ensemble
- current research goals center on ensemble methodology
  - focus on scoring rules, decision theory, forecast-policy alignment

:::

::: {.column width="50%"}

![](Reichlab.png)

![](C19FH.png)

:::

:::

## Evolution of ID Forecasting Hubs {.nostretch}

![](hub_timeline.png){width=80%}

::: {.attribution}
Figure credits: Alex Vespignani and Nicole Samay
:::

New directions: "Hubverse" infrastructure project [https://hubdocs.readthedocs.io/](), genomic targets

::: notes
West Nile mosq control decisions
:::

## Basic products of ID Forecasting Hubs

![](plots/p_VA_CU_IHME.png)

## The Role of ID forecasting Hubs

Outbreak forecasting hubs developed to **"inform public health responses"**

- trend from point and probability (of bins) toward quantile reporting

- uncertainty quantification essential to a public health **DM** (decision maker)

- substantial buy-in from policy makers

> It's important to note that while these efforts represent significant advances in the field, forecasting the spread and impact of infectious diseases is a very complex task with many challenges, and these models should be **interpreted with caution**.

Basic example of **interpretation** is an **allocation decision** regarding

-   medical supplies
-   facility capacity
-   personnel
-   R&D funding
-   public tolerance for NPIs

::: notes
quantiles flexible interpretable - closer to describing full distribution which DM probably always wants

multicenter effort to collect forecasts via hubs for ID outbreaks (epi/pan/endemic)

With uncertainty quantification being well-established as needed DM input,

Allocations can be wrt space, time, demographic

Next: formalize this.
:::

## Scoring rules and social utility

*Very* briefly, a **scoring rule** $S$ assigns a number $S(F,y)$ to the forecast $F$ of $Y$ when $Y=y$ is realized.

$S$ if **proper** if $\displaystyle{E_{F}[S(F,Y)] \leq E_{F}S(G,Y)]}$ for all $F, G$

- ensures that forecasters must give honest forecasts to maximize expected score

Hubs have tried to use proper SRs to rank and combine forecasts in ways that offer best suppport for PH response to ID outbreaks

Primary scoring rule used currently is the **Weighted Interval Score** which is a discrete approximation to the 
**Continuous Ranked Probability Score** (a probabilistic generalization of MAE)

- adopted for convenient scoring of quantile forecasts.

<br>

#### A central goal in design of scoring rules for social good

::: box

Create scenario where the forecaster believes maximizing her own expected utility aligns 
with leading the DM toward actions that maximize social utility.

:::

## Bayes Acts

Scoring rules can be cast in decision-theoretic framework of **Bayes acts**.

<br>

::: box

The **Bayes act** relative to a forecast $Y \sim F$ is the action $x^F$ 
by which a DM will minimize **expected loss** $E_Y[l]$ 
assuming that in reality $Y \sim F$.

:::

 - loss $l$ is a function of the outcome $y$ **and** the action $x$ DM takes in anticipation of $y$

::: box

Key point: Given $l$, we have an *automatically proper* scoring rule 
$S(F,y) := l(x^F, y)$

:::

<br>

#### Our basic example:

 - $x$ and $y$ are vectors in $\mathbb{N}_0^{52}$
 - $y =$ number of severe cases in US states and territories
 - $l(x,y) =$ unmet need when $x < y$ beds are allocated and $y$ severe cases occur

## Formally

$$
\begin{align}
l(x,y) &= \sum_{i=1}^{52} \max(y_i - x_i, 0) = \sum (y_i - x_i)_{+}\\
x^F &\in \left\{ \text{possible } x' \,\middle| \, E_{F}[l(x',Y)] \leq E_{F}[l(x,Y)] \text{ for any possible } x \right\}
\end{align}
$$

<br>

#### Novelty of our approach for ID forecast evaluation:

::: box

Define set of **possible** $x$ as those actions which satisfy $\displaystyle{\sum x_i \leq K}$

:::

<br>

#### Note:

Our loss function and (single) constraint are very simple prototypes meant 
for proof of concept.

::: notes
social/political/economic/psychological/moral utility

this must somehow acknowledge risk attitude heterogeneity among various DM's that a forecast hub might serve

gpl parameters, weights, multiple constraints
:::

## A Lagrangian formulation for the Bayes act

Given $F = \{F_i\}_{i=1}^{N}$, obtaining $x^F$ is a constrained stochastic optimization problem

\begin{align}
L(x, \lambda ; K, F) &= \sum_{i=1}^N E_{F_i}\left[l(x_i, Y_i)]\right)+\lambda\left(\sum_{i=1}^N x_i-K\right) \\

\nabla_{x} L(x^F, \lambda) = 0 &\implies 
E_F \left[\frac{d}{dx_i}(Y_i - x_i^F)_{+}\right] + \lambda \\
&= F_i(x_i^F)-1 + \lambda = 0 
\implies x_i^F = F_i^{-1}(1-\lambda) \\

\nabla_{\lambda} L(x^F, \lambda^{\star}) = 0 
&\implies \sum_{i=1}^{N} F_{i}^{-1}(1-\lambda^{\star}) - K  = 0
\implies 
\boxed{x_{i}^F = F_i^{-1}(1-\lambda^{\star})}
\end{align}

Our allocation scoring rule is then given by the recipe:

1. Use forecast $F$ to form Lagrangian $L$, for expected loss under a resource constraint

2. Solve for optimizer $x^F$ using $L$

3. Plug $x^F$ and outcome $y$ into loss function to get $S(F,y) = l(x^F,y)$

## Basic computational approach

But getting $x^F$ not so simple.

Even for simple parametric $F_i$, there is rarely an analytic solution for $\lambda^{\star}$.

- and we need to reconstruct $F_i$'s from quantile forecasts $q_{i,j}, j = 1,\ldots,23$ to form $L(x, \lambda ; K, F)$

![](plots/dist_pqs_plot_OH.png)

## Iterative methods

So we resort to a binary search for $\lambda^{\star}$ following long history in inventory management of solving **constrained multi-product newsvendor** 
problems^[See e.g., Zhang, Bin, Xiaoyan Xu, and Zhongsheng Hua. 2009. “A Binary Solution Method for the Multi-Product
Newsboy Problem with Budget Constraint.”]

![](plots/iter_combo.png)

## Application

December 2021: Omicron wave clearly started US but forecast teams unsure of severity given uncertainty about $R_0$, cross-protection by vaccination, previous infection, etc.

![](plots/nat_hosps.png)

## Results

::: columns

::: {.column width="30%"}

After adjusting for an oracle the score of an oracle that allocates *fairly* based on what demands actually occur, scores resemble "value" plots from meteorology

- also the Murphy curves of Gneiting et. al.

- extreme shortages or surplusses diminish the oracle's advantage

- ranking at peak seems to be consistent with that in large $K$
region.

:::

::: {.column width="70%"}

![](plots/peak_alloscore.png)

:::

:::

::: aside
Ehm, Werner, Tilmann Gneiting, Alexander Jordan, and Fabian Krüger. 2016. “Of Quantiles and Expectiles: Consistent Scoring Functions, Choquet Representations and Forecast Rankings.”
:::

::: notes
Note: The oracle is not the DGP, it is a time-traveller
:::

## Alloscore and WIS rank models differently

![](plots/p_wis_v_alloscore.png)

## Explanations?

![](plots/mvbucky_dist_alloc.png)

## Limitations

This is **post-hoc** analysis

Hub forecasters were unaware of

- an allocation score (on joint forecast)
- any allocation based loss
- our quantile interpolation/extrapolation methods (`distfromq`)

  - might be especially important for tails

We hope/think that allocation scoring is sensitive to implicit dependence structures in forecasts, but all work so far only refers directly to marginals
- nothing yet with copulas, etc.

<br>

### Thank you!

A very rough R package I wrote to implement scoring procedures:

[https://github.com/aaronger/alloscore](https://github.com/aaronger/alloscore)

A less rough package Evan wrote to implement cdf reconstruction

[https://github.com/reichlab/distfromq](https://github.com/reichlab/distfromq)

<!-- ## Loss vs Utility

Change in loss but not utility from change in - variant severity (assuming target is infections and decision is number of beds)

Change in utility but not in loss from change in - supply and personnel cost structures - public opinion regarding importance of fatality prevention - knowledge of demographic disparities in consequences of outcomes -->

## Data workflow

```
# A tibble: 12 × 9
    time model forecast         ys            K xdf              score_raw score_oracle score
   <int> <chr> <list>           <list>    <dbl> <list>               <dbl>        <dbl> <dbl>
 1     1 m1    <tibble [4 × 7]> <dbl [4]>   400 <tibble [4 × 8]>      62.5          0    62.5
 2     1 m1    <tibble [4 × 7]> <dbl [4]>   450 <tibble [4 × 8]>      50.0          0    50.0
 3     2 m1    <tibble [4 × 7]> <dbl [4]>   400 <tibble [4 × 8]>     150.          27.3 122. 
 4     2 m1    <tibble [4 × 7]> <dbl [4]>   450 <tibble [4 × 8]>     137.           0   137. 
 5     3 m1    <tibble [4 × 7]> <dbl [4]>   400 <tibble [4 × 8]>      62.2          0    62.2
 6     3 m1    <tibble [4 × 7]> <dbl [4]>   450 <tibble [4 × 8]>      49.7          0    49.7
 7     1 m2    <tibble [4 × 7]> <dbl [4]>   400 <tibble [4 × 8]>      54.8          0    54.8
 8     1 m2    <tibble [4 × 7]> <dbl [4]>   450 <tibble [4 × 8]>      40.2          0    40.2
 9     2 m2    <tibble [4 × 7]> <dbl [4]>   400 <tibble [4 × 8]>     157.          27.3 130. 
10     2 m2    <tibble [4 × 7]> <dbl [4]>   450 <tibble [4 × 8]>     149.           0   149. 
11     3 m2    <tibble [4 × 7]> <dbl [4]>   400 <tibble [4 × 8]>      74.9          0    74.9
12     3 m2    <tibble [4 × 7]> <dbl [4]>   450 <tibble [4 × 8]>      70.9          0    70.9
```

```
> a_series_slim$xdf[[9]]
# A tibble: 4 × 8
  target_names     x score_fun     y oracle components_raw components_oracle components
  <chr>        <dbl> <list>    <dbl>  <dbl>          <dbl>             <dbl>      <dbl>
1 A             80.4 <fn>       55.0   51.5             0               3.52      -3.52
2 B             92.2 <fn>      250.   234.            157.             16.0      141.  
3 C            106.  <fn>       89.2   83.5             0               5.70      -5.70
4 D            122.  <fn>       33.5   31.3             0               2.14      -2.14
```

## National picture

![](plots/state_hosps_ens_only.jpeg){width=100%}

## Sharing Business

A Scoring Rule attempts to "share" DM loss with the forecaster through $S(F,y)$

- idea goes back to Savage

- may be in units of money, goodwill, prestige

<br>

<br>

An explicit vacabulary of social loss in ID forecast evaluation has instrinsic advantage
over "off-the-shelf" metrics from statistic inference

::: notes
$S$ a penalty for us

I do not have a clean definition of "social loss/good/risk/regret" but wanted to use term becasue it is in the session title

inference has own decision theory and ideas of utility, loss, Bayes acts
that are related but distinct from the fundamentally econommic/policy focussed
ones we use here
:::