### 2023-10-25 talks/midas2023/gerding - forecast evaluation via social utility.qmd
slides for MIDAS talk

---
title: |
  | Infectious Disease Forecast Evaluation Via Social Utility: Allocation Scores
subtitle: MIDAS 2023
author: Aaron Gerding, Nick Reich, Ben Rogers, Evan Ray
institute: "School of Public Health and Health Sciences, UMass, Amherst"
format: 
  beamer:
    navigation: horizontal
    fontsize: 9pt
    header-includes: |
      \titlegraphic{
      \begin{minipage}{0.4\linewidth}
          \includegraphics[width=\linewidth]{reichlab.png}
      \end{minipage}%
      \hspace{0.06\linewidth}
      \begin{minipage}{0.4\linewidth}
          \includegraphics[width=\linewidth]{C19FH.png}
      \end{minipage}%
      }
  # revealjs: 
  #   width: 1050
  #   height: 700
# revealjs-plugins:
#   - attribution
editor: visual
# css: styles.css
---

## The Role of ID forecasting Hubs

Outbreak forecasting hubs developed to **"inform public health responses"**

Basic examples of **responses** are **allocation decisions** regarding

-   medical supplies
-   facility capacity
-   personnel
-   R&D funding
-   public tolerance for NPIs

## Basic Hub Products

![](plots/p_VA_CU_IHME.png)

::: notes
quantiles flexible interpretable - closer to describing full distribution which DM probably always wants

multicenter effort to collect forecasts via hubs for ID outbreaks (epi/pan/endemic)

With uncertainty quantification being well-established as needed DM input,

Allocations can be wrt space, time, demographic

Next: formalize this.
:::

## Scoring rules and social welfare

A **scoring rule** $S$ assigns a number $S(F,y)$ to the forecast $F$ of $Y$ when $Y=y$.

$S$ if **proper** if $\displaystyle{E_{F}[S(F,Y)] \leq E_{F}S(G,Y)]}$ for all $F, G$

-   ensures that honesty maximizes expected score

Hub goal of increasing social welfare by ranking and combining forecasts to best support PH responses seems to require use of proper scores.

Current standard is the **Weighted Interval Score** (a discrete version of the **Continuous Ranked Probability Score**)

-   adopted for convenient scoring of quantile forecasts.

<br>

#### A central goal in design of scoring rules for social welfare

::: box
Create scenario where the forecaster believes maximizing her own expected utility aligns with leading the DM toward actions that maximize social utility.
:::

## Bayes Acts

Scoring rules can be cast in decision-theoretic framework of **Bayes acts**.

::: box
The **Bayes act** relative to a forecast $Y \sim F$ is the action $x^F$ by which a DM will minimize **expected loss** $E_F[l(x,Y)]$.
:::

-   loss $l(x,y)$ is the **consequence** of experiencing $y$ having prepared with $x$

::: box
Key point: $S(F,y) := l(x^F, y)$ is an *automatically proper* scoring rule.
:::

<br>

#### Our basic example:

-   $x$ and $y$ are vectors in $\mathbb{N}_0^{52}$
-   $y =$ number of severe cases in US states and territories
-   $l(x,y) =$ unmet need when $x$ beds allocated and $y$ severe cases occur

```{=tex}
\begin{align*}
l(x,y) &= \sum_{i=1}^{52} \max(y_i - x_i, 0)\\
x^F &= \text{minimizer of } E_{F}[l(x,Y)] \text{ over all feasible } x
\end{align*} 
```
<br>

**Our new idea:** Define **feasible** $x$ as satisfying $\displaystyle{\sum x_i \leq K}$

<br>

::: notes
social/political/economic/psychological/moral utility

this must somehow acknowledge risk attitude heterogeneity among various DM's that a forecast hub might serve

gpl parameters, weights, multiple constraints
:::

## Computation

Obtaining $x^F$ is a constrained stochastic optimization problem

-   known in inventory management as a **constrained multi-product newsvendor** problem

-   formally solvable using Lagrange multiplier method to get a quantile representation $x_i^F = F_i^{-1}(\tau(K,F)), i = 1,\ldots,N$

-   in practice, we find $x_i^F$'s via an iterative method:

![](plots/iter_combo.png)

## Application

December 2021: Omicron wave clearly started US but forecast teams unsure of severity given uncertainty about $R_0$, cross-protection by vaccination, previous infection, etc.

![](plots/nat_hosps.png)

## Oracle adjusted allocation scores near Omicron peak

![](plots/peak_alloscore.png)

## Some Observations

::: columns
::: {.column width="40%"}
-   related to the Murphy curves of Ehm, Gneiting, Jordan, Krüger, 2016

-   extreme shortage or surplus diminishes oracle's advantage

-   ranking consistent across large $K$ region.
:::

::: {.column width="60%"}
![](plots/peak_alloscore_min.png)
:::
:::

## Alloscore and WIS rank models differently

![](plots/p_wis_v_alloscore.png)

## Explanations?

![](plots/mvbucky_dist_alloc.png)

## Limitations

This is **post-hoc** analysis

Hub forecasters were unaware of

-   an allocation score (on joint forecast)

-   any allocation based loss

-   our quantile interpolation/extrapolation methods (`distfromq`)

    -   might be especially important for tails

We hope/think that allocation scoring is sensitive to implicit dependence structures in forecasts, but all work so far only refers directly to marginals - nothing yet with copulas, etc.

<br>

### Thank you!

A very rough R package I wrote to implement scoring procedures:

<https://github.com/aaronger/alloscore>

A less rough package Evan wrote to implement cdf reconstruction

<https://github.com/reichlab/distfromq>