### 2023-11-03 talks/midas2023/gerding - forecast evaluation via social utility.qmd
separate plotting code into talks/plots/

---
title: |
  | Infectious Disease Forecast Evaluation Via Social Utility: Allocation Scores
subtitle: MIDAS 2023
author: Aaron Gerding, Nick Reich, Ben Rogers, Evan Ray
institute: |
 | School of Public Health and Health Sciences, UMass, Amherst
 | Department of Biostatistics and Epidemiology
format: 
  beamer:
    keep-tex: true
    navigation: horizontal
    fontsize: 9pt
    header-includes: |
      \titlegraphic{
      \begin{minipage}{0.4\linewidth}
          \includegraphics[width=\linewidth]{reichlab.png}
      \end{minipage}%
      \hspace{0.06\linewidth}
      \begin{minipage}{0.4\linewidth}
          \includegraphics[width=\linewidth]{C19FH.png}
      \end{minipage}%
      }
  # revealjs: 
  #   width: 1050
  #   height: 700
# revealjs-plugins:
#   - attribution
editor: visual
# css: styles.css
---

## Outbreak Forecast Hubs and **Informed** Resource Allocation

Hubs developed to **"inform public health responses"**

-   such as how resources are **allocated** among locations

-   e.g., medical supplies, facility capacity, personnel

![](../plots/output/p_VA_JHU_MUNI.jpeg)

::: notes
quantiles flexible interpretable - closer to describing full distribution which DM probably always wants

multicenter effort to collect forecasts via hubs for ID outbreaks (epi/pan/endemic)

With uncertainty quantification being well-established as needed DM input,

Allocations can be wrt space, time, demographic

Next: formalize this.
:::

## Scoring rules and social welfare

Hubs strive to rank and combine forecasts so as to optimize social welfare via the public health decisions that forecasts inform

-   uncertainty quantification, and therefore, probabilistic forecasting essential

-   basic strategy: use a **scoring rule** $S$ which assigns a loss $S(F,y)$ when a probabilistic forecast $F$ of $Y$ is chosen and $Y=y$ is observed.

-   optimizing welfare requires that forecasters say what they believe; so $S$ should be **proper** meaning $\displaystyle{E_{F}[S(F,Y)] \leq E_{F}S(G,Y)]}$ for all $F, G$

Current standard is the **Weighted Interval Score** (discrete CRPS)

-   adopted largely for convenient scoring of quantile forecasts.

## Tools from decision theory

### A central goal in design of scoring rules

Tie forecast scores directly to the benefits to society of the decisions they inform

Key tools from decision theory for linking success/failure of forecast-informed policy actions to scoring rules:

-   Let $l(x,y)$ be the **loss** of experiencing $y$ after taking policy action $x$

-   The **Bayes risk** of a forecast $Y \sim F$ is $\min_x E_F[l(x,Y)]$

-   A **Bayes act** for $F$ is an action $x^F$ that attains this minimum

-   Losses from Bayes acts define an *automatically proper* scoring rule

$$
S(F,y) := l(x^F, y)
$$

**Note:** Scoring rules are really only a means to an end.  It is the expected score, in this case the Bayes risk, which characterize the value a forecast adds for a decision maker.  Scoring rule sample averages estimate the expected score.

## A new scoring rule via a new loss function for constrained actions

### Our basic example:

-   $x$ and $y$ are vectors in $\mathbb{N}_0^{52}$

-   $y =$ number of severe cases in US states and territories

-   $l(x,y) =$ unmet need when $x$ beds allocated and $y$ severe cases occur

```{=tex}
\begin{align*}
l(x,y) &= \sum_{i=1}^{52} \max(y_i - x_i, 0)\\
x^F &= \text{minimizer of } E_{F}[l(x,Y)] \text{ over all feasible } x
\end{align*} 
```
<br>

**Our new idea:** Define **feasible** $x$ as satisfying $\displaystyle{\sum x_i \leq K}$

<br>

::: notes
social/political/economic/psychological/moral utility

this must somehow acknowledge risk attitude heterogeneity among various DM's that a forecast hub might serve

gpl parameters, weights, multiple constraints
:::

## Computation

Obtaining $x^F$ is a constrained stochastic optimization problem

-   known in inventory management as a **constrained multi-product newsvendor** problem

-   formally solvable using Lagrange multiplier method to get a quantile representation $x_i^F = F_i^{-1}(\tau(K,F)), i = 1,\ldots,N$

-   in practice, we find $x_i^F$'s via an iterative method:

![](../plots/output/iter_combo.png)

## Application

December 2021: Omicron wave clearly started US but forecast teams unsure of severity given uncertainty about $R_0$, cross-protection by vaccination, previous infection, etc.

![](../plots/output/nat_hosps.png)

## Oracle adjusted allocation scores near Omicron peak

![](../plots/output/peak_alloscore.png)

## Some Observations

::: columns
::: {.column width="40%"}
-   related to the Murphy curves of Ehm, Gneiting, Jordan, Krüger, 2016

-   extreme shortage or surplus diminishes oracle's advantage

-   ranking consistent across large $K$ region.
:::

::: {.column width="60%"}
![](../plots/output/peak_alloscore_min.png)
:::
:::

## Alloscore and WIS rank models differently

![](../plots/output/p_wis_v_alloscore.png)

## Explanations?

![](../plots/output/mvbucky_dist_alloc.png)

## Limitations

This is **post-hoc** analysis

Hub forecasters were unaware of

-   an allocation score (on joint forecast)

-   any allocation based loss

-   our quantile interpolation/extrapolation methods (`distfromq`)

    -   might be especially important for tails

We hope/think that allocation scoring is sensitive to implicit dependence structures in forecasts, but all work so far only refers directly to marginals - nothing yet with copulas, etc.

<br>

### Thank you!

A very rough R package I wrote to implement scoring procedures:

<https://github.com/aaronger/alloscore>

A less rough package Evan wrote to implement cdf reconstruction

<https://github.com/reichlab/distfromq>