### 2024-02-14 alloscore_manuscript/alloscore-application-shorter.tex

\documentclass{article}\usepackage[]{graphicx}\usepackage[]{xcolor}
% maxwidth is the original width if it is less than linewidth
% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

\usepackage{amsmath, amsfonts, amssymb}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{cases}
\usepackage{caption}
\usepackage{xr-hyper}
\usepackage{hyperref}
\usepackage{xurl}
\externaldocument{supplement_short}

\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator{\short}{sh}
\DeclareMathOperator{\Ex}{\mathbb{E}}

\usepackage{setspace}
\onehalfspacing

\usepackage{parskip}

\usepackage{soul}
\usepackage{xcolor}
\def\elr#1{{\color{cyan}\textbf{ELR:[#1]}}}
\def\apg#1{{\color{red}\textbf{APG:[#1]}}}
\def\bwr#1{{\color{violet}\textbf{BWR:[#1]}}}
\def\ngr#1{{\color{blue}\textbf{NGR:[#1]}}}

\usepackage{natbib}
\bibliographystyle{unsrtnat-abbr}

\title{Evaluating infectious disease forecasts \\ with allocation scoring rules}
\author{Aaron Gerding, Nicholas G. Reich, Benjamin Rogers, Evan L. Ray}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

\newcommand{\del}[2]{\frac{\partial {#1} }{\partial {#2}} }
\newcommand{\dby}[2]{\frac{d {#1} }{d {#2}} }
\newcommand{\sbar}{\overline{s}}
\newtheorem{proposition}{Proposition}

\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\maketitle

\begin{abstract}

Recent years have seen increasing efforts to forecast infectious disease burdens, with a primary goal being to help
public health workers make informed policy decisions. However, there has only been limited discussion of how
predominant forecast evaluation metrics might indicate the success of policies based in part on those forecasts. We
explore one possible tether between forecasts and policy: the allocation of limited medical resources so as to minimize
unmet need. We use probabilistic forecasts of disease burden in each of several regions to determine optimal resource
allocations, and then we score forecasts according to how much unmet need their associated allocations would have
allowed. We illustrate with forecasts of COVID-19 hospitalizations in the US, and we find that the forecast skill
ranking given by this allocation scoring rule can vary substantially from the ranking given by the weighted interval
score. We see this as evidence that the allocation scoring rule detects forecast value that is missed by traditional
accuracy measures and that the general strategy of designing scoring rules that are directly linked to policy
performance is a promising direction for epidemic forecast evaluation.

\end{abstract}

\section{Introduction}

Infectious disease forecasting models have emerged as important tools in public health outbreak response. The
predictions they provide increasingly inform decisions regarding a wide variety of countermeasures intended to reduce
transmission and mitigate the severity of disease outcomes. For example, estimates of the onset time of the flu season
have been used in developing national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola and
diptheria dynamics have been made with the clearly stated goal of helping local public health workers choose the timing
and location of interventions in settings where resources are severely constrained 
\citep{meltzer2014estimating, rainisch2015regional, camacho2015-ebola-bed,finger_real-time_2019}. More recently, in the
context of the outbreak of COVID-19 across the US, \cite{bertsimas2021predictionsCOVID} used forecasts as inputs to
decision tools for the interstate reallocation of ventilators and ICU capacity, and to recommend vaccine trial sites to
a major trial sponsor. \cite{fox_real-time_2022} similarly used predictive models to inform intrastate resource and
care site planning, as well as local community guidelines for masking, traveling, dining and shopping (University of
Texas, 2022)\nocite{utnews2022}.

In the wake of the COVID-19 pandemic, this trend has been followed by calls for infectious disease forecasts to be not
only designed, but also evaluated in ways that align specifically with how forecasts can be used to inform such outbreak
control decisions \citep{marshall2023predictions, bilinski_adaptive_2023}. This contrasts, however, with the
historically standard practice of measuring the quality of disease forecasts using general purpose accuracy and skill
scores, especially those that have implementations available in existing software when the relevant outbreak occurs. For
point forecasts, the root mean square error (RMSE) (e.g., \cite{papastefanopoulos2020covid}) and the mean absolute error
(MAE) (e.g., \cite{johansson2016evaluating}) are common choices. For probabilistic forecasts, which are the focus of
this paper, researchers have often relied on the logarithmic score (LS). For example, the LS has been used to evaluate
the skill of US seasonal influenza forecasts \citep{mcgowan_collaborative_2019,reich_collaborative_2019} as well as
forecasts targeting surveillance measures of dengue incidence in Peru and Puerto Rico \citep{johansson_open_2019}. More
recently, the continuous ranked probability score (CRPS) and a discretized version adapted to multi-quantile forecasts,
the weighted interval score (WIS) \citep{bracher2021evaluating}, have gained prominence. For example, the CRPS was used
to assess probabilistic forecasts (based on random effect models) of dengue incidence at the district level in Vietnam
\citep{colon-gonzalez_probabilistic_2021}. And the WIS has been used during the COVID-19 pandemic to evaluate forecasts
of observed cases, hospitalizations and deaths in the US and Europe, as reported by municipal, state, and federal
surveillance systems \citep{cramer_evaluation_2022,fox_real-time_2022,sherratt2023predictive}.

While it should be noted that there are ways to interpret any of these scores abstractly through the lens of decision
theory, and that all of the application-specific papers cited above benefited from direct collaboration with public
health agencies, a key impetus for the present work has been that we were not able to find in any of them, nor in the
literature they represent, explicit connections between how a forecast was evaluated and how that forecast was used in
practice.

A general phenomenon at play here --- one that has been observed repeatedly over the past few decades in other fields
such as finance, supply chain management, and meteorology --- is that while scores such as RMSE, MAE, LS, CRPS, and WIS
can describe the \emph{quality} of a forecast in terms of how well it corresponds to the observed disease outcome, they
will often fail to register the \emph{value} of a forecast in the context of a specific decision. We refer the reader to
\cite{yardley2021utility_cost_forecasts} and the references collected therein (especially the foundational
\cite{murphy1993whatisagoodforecast}) for a general overview touching on a wide range of forecasting contexts and  also
to \cite{pesaran2002decision_based_eval} for a clear discussion from an econometric perspective.

Despite this now well-developed discussion of the quality-value distinction in the larger forecasting community, we are
aware of only a limited literature attempting to connect the value of \emph{infectious disease} forecasts to their
impact on and through policy.  And within this body of work we have found the discussion of such a connection to still
be at a formally and quantitively imprecise stage. In \cite{ioannidis2022forecastingCOVIDfailed}, the possible negative
consequences of inaccurate forecasts of infectious disease are discussed, but there is no attempt to quantify the
utility or loss incurred as a result of those forecasts. \cite{bilinski_adaptive_2023} explore ways in which predictive
classifiers of local COVID-19 risk levels in the US could be tuned to policymaker preferences for different costs
associated with over- and under-reaction to disease dynamics, but they do not clearly identify the source of these
costs or how they depend on quantifiable policy choices. A similar discussion related to dengue countermeasures in
Vietnam appears in \cite{colon-gonzalez_probabilistic_2021}.  A novel version of the WIS informally motivated by
utility considerations is developed in \cite{marshall2023predictions}, but the score is not derived in a
decision-theoretic manner.  There is also a thread of literature that frames infectious disease modeling as a component
of a larger system for understanding how policy goals, means, and choices interact and constrain one another. As an
example, \cite{Probert2016decisionMakingFootMouth} explore how policy recommendations ought to flow from a possibly 
incongruous set of simulation-based projection models of a hypothetical foot-and-mouth disease outbreak when there are 
ranges of plausible responses and stakeholder interests. Decision theory plays a prominent role here, but not explicitly
as a way to evaluate the choices made in developing the models.

In this work, we begin to fill this gap between the ways that infectious disease forecasts have traditionally been
evaluated and the ways that they have been used to support public health policy. To do so, we consider a setting in
which forecasts are used to help determine the allocation of a limited quantity of medical supplies across multiple
regions. In section~\ref{sec:methods} of the paper, we define a new forecast scoring rule --- the {\em allocation score}
--- that evaluates forecasts based on how beneficial resource allocations derived from them would turn out to be. In
section \ref{sec:application}, we present an illustrative analysis using the allocation score to evaluate forecasts of
hospital admissions in the US that were made leading up to and during the Omicron wave that peaked in January of 2022.
This analysis is ``synthetic'' insofar as it is not intended to correspond to any specific historical record of
allocation decisions that could have been supported by hospitalization forecasts during this period. However, we view
the general allocation problem on which our framework is based as a versatile template for formalizing real-world
decisions that must constantly be made in real-time by public health administrators around the globe, especially those
related to hospital capacity, ventilator usage, doses needed for specific medications and other situations where an
outbreak creates sudden and highly variable demand for potentially scarce resources.

\section{The Allocation Score}
\label{sec:methods}

We begin with an informal description of the allocation score and some examples illustrating its key characteristics in
section \ref{sec:methods.overview}. In section \ref{sec:methods.detailed} we develop the allocation score more
carefully, using a decision theoretic procedure for deriving proper scoring rules. (See section \ref{sec:proper} of the
supplement for a definition of a proper scoring rule and a more technical discussion of the procedure). In section
\ref{sec:methods.related}, we note that another group of common scores including the quantile score, WIS, and CRPS, can
also be derived from decision theoretic foundations \textemdash starting from a different decision making context.

\subsection{Overview of Allocation Scoring}
\label{sec:methods.overview}

Suppose that a decision maker is tasked with determining how to allocate $K$ available units of a resource across $N$
locations. If the decision maker is provided with a multivariate forecast $F$ where each marginal forecast distribution
$F_i$ predicts resource need in a particular location, one option is to choose the resource allocation that minimizes
the expected total unmet need according to the forecast. We will give a more precise mathematical statement in section
\ref{sec:methods.detailed}, but informally, the total expected unmet need according to the forecast is
\begin{align}
\sum_{i=1}^N \mathbb{E}_{F_i}[\text{unmet need in location $i$}], \label{eqn:informal_objective}
\end{align}
where the unmet need in a particular location is the difference between resource need in that location and the number of
resource units that were allocated there. This allocation problem has an intuitively appealing solution: allocate so that the
probabilities of need exceeding allocation in various locations are as close to each other as possible. This will lead
to the allocations provided by $F$ being quantiles of the marginal distributions $F_i$ for some \emph{single}
probability level $\tau$ that is shared in common for all locations.

After time passes and the actual level of resource need has been observed, the value of a selected allocation can be
measured by comparing the actual need in each location to the amount of resources that were sent there. Specifically, we
compute the total unmet need that resulted from the selected allocation:
\begin{align}
    \sum_{i=1}^N \text{unmet need in location $i$}. \label{eqn:informal_loss}
\end{align}
%We emphasize that in Equation \eqref{eqn:informal_loss} the calculation of unmet need is based on the actual resource need that was realized in each location, while in Equation \eqref{eqn:informal_objective} the calculation of unmet need was based on the forecast distribution of future levels of resource need.
We regard one allocation as better than another (with respect to the realized need) if it results in lower total 
unmet need, and accordingly regard one forecast as better than another (again with respect to the realized need)
if the allocation derived from the former results in lower total unmet need than does the allocation derived from the 
latter.

The \textbf{allocation score} of the forecast $F$ is the avoidable unmet need that results from using the allocation
that minimizes the expected unmet need according to that forecast. By ``avoidable unmet need'', we mean that the
allocation score does not include the amount of unmet need that was inevitable simply because the amount of available
resources $K$ was less than the need for resources. Rather, the allocation score measures the unmet need that could have
been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so
that nothing is wasted in one location while it could be put to use in another. The best and lowest possible allocation 
score is 0.  A positive allocation score indicates some way in which an oracle could have used perfect foreknowledge to
meet more of the need than did the allocation suggested by $F$.

\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$
and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$. The means of these distributions are
given by the scale parameters $\sigma_i$. When the marginal forecasts are exponential distributions, it can be shown
that the optimal allocation divides the available resources among the locations proportionally to the scale parameters
$\sigma_i$ (see section \ref{sec:bayes-quantiles} of the supplemental materials). If $K = 5$ units of the resource are
available, the optimal allocation according to $F$ would be 1 and 4 resource units in locations 1 and 2, respectively. 
Increasing $K$ to $10$ changes this optimal allocation to 2 and 8. Figure~\ref{fig:exp_alloc_example} illustrates the 
situation.

\begin{figure}
    \includegraphics[width=\textwidth]{../figures/exponential_pred_expected_loss.pdf}
    \caption{An illustration of the resource allocation problem in Example 1. There are $N = 2$ locations, with
    predictive distributions $F_1 = \mathrm{Exp}(1)$ and $F_2 = \mathrm{Exp}(1/4)$. The cumulative distribution
    functions of these distributions are illustrated in the panels at bottom and right. In the center panel, the
    background shading corresponds to the expected loss according to these forecasts. Diagonal black lines indicate
    resource constraints at $K=5$ and $K=10$ units; any point along those lines corresponds to an allocation that meets
    the resource constraint. For these forecasts, the optimal allocations are $(1, 4)$ for $K=5$ and $(2, 8)$ for
    $K=10$. These allocations are at the point on the constraint line where the expected loss is smallest, which also
    corresponds to the point where a level set of the expected loss surface (blue curve) is tangent to the constraint.}
    \label{fig:exp_alloc_example}
\end{figure}

Next suppose that we observe resource needs of 1 and 10 in locations 1 and 2, respectively. Based on these observed
needs, we evaluate the allocation suggested by the forecast by calculating the amount of unmet need
that resulted from that allocation over and above what was unavoidable given the resource constraint. With $K = 5$ 
available resource units, the allocation based on the forecast exactly meets the observed need in location 1, but it leaves 6
units of need unmet in location 2. But these 6 units of unmet need are unaviodable under the constraint $K=5$: allocating 0 units of resources to location 1 and 5 to location 2, for example, still results in a total unmet
need of 6 across both locations. This gives an allocation score of 0 for the forecast when $K = 5$. On the other hand, when
$K = 10$, the forecast's allocation results in $10 - 8 = 2$ units of unmet need in location 2 despite leaving no need
unmet in location 1. In this case, the oracle would be able to prevent all but 1 of the total 11 units of need from
going unmet, for example by allocating 1 unit of resources to location 1 and the remaining 9 units of resources to
location 2. So for $K = 10$, the forecast gets an allocaton score of 1 (= 2 realized $-$ 1 unavoidable) in
units of avoidable unmet need.

These scores illustrate a general result: allocation scores for a forecast will tend to be larger when the resource
constraint is close to the observed need, because this is when it matters most which locations are allocated more or
less resources. If the amount of available resources is small relative to the observed need, any allocation of those
limited resources will result in a large amount of unmet need. If the amount of available resources is comparatively
large, it becomes less important which locations receive relatively more or fewer resources because all locations are
likely to
receive enough resources to meet their need. In either of these extremes of resource availability, the avoidable unmet
need that arises from the allocation suggested by a forecast (i.e., the forecast's allocation score) will tend to be
small.

\paragraph{Example 2} Now consider a different forecast that also has exponential distributions for resource need in
each location, but that has the scale parameters $\sigma_1 = 2$ and $\sigma_2 = 8$, twice as large as the scale
parameters of the forecast in Example 1. Because the optimal allocation is proportional to the scale parameters, this
forecast would lead to the same allocations as the forecast in Example 1, and would therefore be assigned the same
allocation score.

Examination of these results leads to two observations. First, the reason that these forecasts had a positive (i.e.,
non-optimal) allocation score at $K=10$ is that they did not get the relative magnitude of resource need across the two
locations right: the realized need was 10 times as large in location 2 as in location 1, but the forecasts only
indicated that the resource allocation for location 2 should be 4 times the allocation for location 1. At its core, the
allocation score measures whether the forecast accurately captures the relative magnitudes of resource need across
different locations, which is precisely the information that is needed to allocate resources to those locations subject
to a fixed resource constraint.

A second observation is that the forecasts in examples 1 and 2 predicted different mean levels of resource needs, but
had the same allocation score. The allocation score does not directly measure whether the forecasts correctly capture
the absolute magnitude of resource need in each individual location. This stands in contrast to other common scoring
methods that aggregate scores such as log score, CRPS, or WIS for each location, where a poor forecast for one location
is penalized regardless of alignments in other units.
%Note that we do not claim that the allocation score is generically preferable to these other scores \textemdash rather, it provides a view of forecast performance that is specifically tuned to the context of decision making about resource allocations.

\subsection{A decision theoretic development of the allocation score}
\label{sec:methods.detailed}

We give a high-level review of a general procedure for developing proper scoring rules that are tailored to specific
decision making tasks in section \ref{sec:methods.detailed.decisiontheory}, and then in section
\ref{sec:methods.detailed.specific_allocation} we apply that procedure to develop the allocation score based on the task
of deciding on how to allocate a fixed supply of resources across multiple locations. In
\ref{sec:methods.detailed.integrated_allocation} we consider a small extension where the resource constraint is not
known, or it is desired to consider the value of forecasts across a range of decision making scenarios. This gives rise
to the \emph{integrated allocation score}.

\subsubsection{The decision theoretic setup for forecast evaluation}
\label{sec:methods.detailed.decisiontheory}

%In this section, we give an overview of the decision theoretic setup for developing proper scoring rules that measure
%the value of a forecast as an input to decision making. We keep the discussion here at a somewhat informal level; we
%refer the reader to [some subset of Brehmer and Gneiting; Grünwald and Dawid; Dawid; Granger and Pesaran 2000; Granger
%and Machina 2006; Ehm et al. 2016] for more technically precise discussion.

Decision theory is concerned with understanding how an \emph{action} $x$ both is and ought to be selected from a set of
possible actions $\mathcal{X}$ --- that is, how a decision is and ought to be made --- based on the anticipated
consequences of $x$ under future states. We represent such future states as a random variable $Y$ and refer to a
realization $y$ of $Y$ as an \emph{outcome}. In this paper $x$ will be related to the level of investment in a measure
designed to mitigate severe disease outcomes such as hospital beds, ventilators, medication, or medical staff, with
$\mathcal{X}$ being the set of all feasible levels of investment, while an outcome $y$ will correspond to the number of
individuals who eventually become sick and would benefit from the mitigation measure. A basic strategy in decision
theory is to assume that the consequences of an action $x$ under outcome $y$ can be assigned a numerical value, or
\emph{loss}, measuring the success of $x$, which in our examples will be the extent to which $x$ meets the realized need
$y$.  One way that a decision maker can try to anticipate the consequence of an action $x$ is by using a forecast $F$ of
$Y$. The loss of the action $x$ chosen according to $F$ can, in turn, be intepreted as the value of $F$ as an input to
the decision making process under the realized outcome $y$ of $Y$.

We arrive through this discussion at a three-step procedure for developing scoring rules for probabilistic forecasts:
\begin{enumerate}
  \item Specify a \emph{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$
    eventually occurs.
  \item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under
    the distribution $F$.
  \item The \emph{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: 
    $S(F, y) = s(x^F, y)$.
\end{enumerate}
This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss
function. We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} and in section
\ref{sec:proper} of the supplement we review the basic fact from the scoring rule literature (e.g.,
\cite{dawid2007geometry,gneiting2007strictly}) that Bayes scoring rules are proper by construction.

\subsubsection{The allocation score for a fixed resource constraint}
\label{sec:methods.detailed.specific_allocation}

In our setting an action is a vector $x = (x_1, \ldots, x_N)$ specifying the amount of resources allocated to each of
$N$ locations. We require that $0 \leq x$, i.e., that each $x_i$ is non-negative, and that the total allocation across
all locations equals the amount of available resources, $K$: $\sum_{i=1}^N x_i = K$. The set $\mathcal{X}$ consists of
all possible allocations that satisfy these constraints. The eventually realized resource need in each location is
denoted by $y = (y_1, \ldots, y_N)$. The future outcome that will determine the loss caused by choosing the allocation
$x$ is the vector of resource needs $y = (y_1, \ldots, y_N)$ for each location which is a realization (unknown at the
time of decision making) of the random vector $Y = (Y_1, \ldots, Y_N)$. A forecast of the random need $Y$ for each
location is written as $F = (F_1, \ldots, F_N)$. We assume need is non-negative and the forecasts reflect that, i.e. the
support of each $F_i$ is a subset of $\mathbb{R}^+$. Finally, we assume that any unmet unit of need in any location
incurs a fixed loss of $L>0$, so that if the selected resource level $x_i$ in location $i$ is less than the realized
need $y_i$, a loss of $L \cdot (y_i - x_i)$ results. A variety of extensions to this setup are possible; for example, we
might account for storage costs for resources that go unused, allow for a different loss per unit of unmet need in each
location, or account for resource transportation costs. In this work, we choose to keep the loss function relatively
simple to focus on the core ideas.

It is helpful to clearly distinguish between the time $t_d$ when a \emph{d}ecision is made about a public health
resource allocation and the time $t_r$ when \emph{r}esource needs that might be addressed by that allocation occur. Our
setup assumes that $t_d < t_r$ and that the resource in question can only meet the need realized at $t_r$, not prevent
it. That is, we stipulate that the random vector $Y$ of need is independent from the chosen allocation $x$. In the
context of infectious disease, this means that we do not consider resources, such as vaccines, that are intended to
reduce the number of people who will become sick at some point in the future. Instead, our setup addresses resources
like hospital beds, oxygen supply, or ventilators which are intended to meet the medical needs of patients who are
already sick. Additionally, our setup addresses decision-making that is related to resource needs only at the time
$t_r$; we do not consider sequences of multiple decisions that are made over time or account for the impact of decisions
on resource needs at any time other than $t_r$. We outline some opportunities to extend our work to more complex
decision making settings in the discussion.

With this problem formulation in place, we can develop a proper scoring rule following the procedure in section
\ref{sec:methods.detailed.decisiontheory}.

\paragraph{Step 1: specify a loss function.} The loss incurred by an allocation $x$ under an outcome $y$ is the sum of
contributions from unmet need in each location:
\begin{equation}
  s_A(x, y) = \sum_{i=1}^N L \cdot \max(0, y_i - x_i). \label{eqn:loss_fn}
\end{equation}
Here, $\max(0, y_i - x_i)$ is the unmet need in location $i$, which is given by $y_i - x_i$ if the realized need $y_i$
in location $i$ is greater than the amount $x_i$ allocated to that location, or $0$ if the amount $x_i$ allocated to
unit $i$ is greater than or equal to the realized need. Also, $L$ is a constant scalar value, the same across all
locations, specifying the ``cost'' of one unit of unmet need.

\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act associated with the
forecast, $x^{F,K}$, is the allocation that minimizes the expected loss, that is, the solution of the \emph{allocation
problem} associated with $K$:
\begin{align}
  \underset{0 \leq x}{\mathrm{minimize}}\,\, \mathbb{E}_{F} [s_A(x, Y)] \text{ subject to }
  \, \sum_{i=1}^N x_i = K, \label{AP}
\end{align}
where $\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[\max(0, Y_i - x_i)]$ sums the expected loss
due to unmet need across all locations.

In the supplement we show that the components of the Bayes act are quantiles $x_i^{F,K} = F_i^{-1}(\tau^{F,K})$ at a
probability level $\tau^{F,K}$ that depends on the forecast $F$ and the resource constraint $K$, but is shared across
all locations. This probability level is the level at which the resource constraint is satisfied: $\sum_{i=1}^N
F_i^{-1}(\tau^{F,K}) = K$. This tells us that in order to allocate optimally (according to $F$), we must divide
resources among the locations so that there is an equal forecasted probability in every location that the allocation is
sufficient to meet resource need. This solution to the allocation problem is well-known in inventory management and is
often attributed to \cite{hadleywhitin1963}.

\paragraph{Step 3: Define the scoring rule.} We can now use the Bayes act to define a proper scoring rule for the
probabilistic forecast $F$. Consider first the ``raw'' score defined as
\begin{align}
  S_A^{\text{raw}}(F, y; K) = s_A(x^{F,K}, y) = \sum_{i=1}^N L \cdot \max(0, y_i - x_i^{F,K}).
\end{align}
This measures the total unmet need across all locations that results from using the Bayes allocation associated with the
forecast $F$ when the actual level of need in each location is observed to be $y_i$.

To make this a more easily interpreted measure of forecast performance, we will adjust the raw score by subtracting the
minimum loss achievable by an \emph{oracle} allocator which has precise foreknowledge of the outcomes $y_i$. When the
oracle has sufficient resources to meet the total need, i.e., when $\sum_{i=1}^{N}y_i \leq K$, the oracle's loss is zero
and allocation score coincides with the raw score. On the other hand, when the oracle cannot cover all need and incurs a
loss of $L \cdot \left(\sum_{i=1}^{N}y_i - K \right) > 0$, we adjust the raw score by this loss.
The oracle-adjusted score can therefore be written as
\begin{align}
  S_A(F, y; K)  &= S_A^{\text{raw}}(F, y; K) - L \cdot \max\left(0,\sum_{i=1}^{N}y_i - K\right) \\
  &= L\left\{\sum_{i=1}^N \max(0, y_i - x_i^{F,K}) -  \max\left(0,\sum_{i=1}^{N}y_i - K\right)\right\}.
\end{align}
The oracle adjustment aligns with a common theme in economic decision theory that \emph{opportunity loss} (often known
as \emph{regret} or (negative) \emph{relative utility}) is often a more important quantity than absolute loss (see e.g.,
\cite{DIECIDUE201788}).

\subsubsection{Integrating the allocation score across resource constraint levels}{}
\label{sec:methods.detailed.integrated_allocation}

The allocation score $S_A$ developed in the previous section evaluates the forecast distributions
$F$ based on a single probability level $\tau^{F,K}$. This is appropriate if the resource constraint $K$ is a known
constant. However, if $K$ is not precisely known at the time of decision making or there is interest in measuring the
value of forecasts across a range of decision making scenarios with different resource constraints, we can use an
\emph{integrated allocation score} (IAS) that integrates the allocation score across values of $K$, weighting by a
distribution $p$:
$$S_{IAS}(F, y) = \int S_A(F,y; K) p(K) \, dK$$
We note that the device of considering a range of hypothetical decision makers or decision making problems with
different problem parameters has been employed in the past \cite[e.g.,][]{murphy1993whatisagoodforecast}.

\subsection{Connections to Other Scores}
\label{sec:methods.related}

The weighted interval score (WIS) was proposed in 2020 as a way to score forecasts that were being made in the early
stages of the COVID-19 pandemic \citep{bracher2021evaluating}; equivalent scores had also been used in previous forecast
evaluation efforts \cite[e.g.,][]{hong2016probabilisticEnergyForecasting}. The WIS is a proper scoring rule for
forecasts that use a set of quantiles to represent a probabilistic forecast distribution. While pointing a reader
interested in more mathematical detail to \cite{bracher2021evaluating}, we note simply that the WIS is a weighted sum of
interval scores at different probability levels (e.g., 50\% prediction intervals, 80\% PIs, 95\% PIs, etc...). Larger
interval scores indicate less skillful forecasts. An interval score consists of (a) the width of the interval, with
larger intervals receiving higher scores (higher scores indicate less accuracy), and (b) a penalty if the interval does
not cover the eventual observation, which increases the further away the interval is from the observed value.
Equivalently, the WIS can also be characterized as a weighted sum of quantile scores for each individual predictive
quantile. The quantile score for a particular quantile level assigns an asymmetric penalty to predictions that are too
high or too low, with the relative sizes of the penalties set so that in expectation the score is minimized by the given
quantile of the distribution. The most commonly used version of WIS is one that uses an equal weighting of all quantile
levels, in which case WIS approximates the continuous ranked probability score (CRPS), a commonly used score for
probabilistic forecasts. \emph{It is important to note that this weighting was proposed because the resulting score
approximates the CRPS, and not because it aligned with any particular public health decision-making rationale.}

That said, the quantile score and WIS can be derived using the same decision theoretic procedure that we outlined in
section \ref{sec:methods.detailed}. In fields such as meteorology and supply chain management, a great deal of attention
has been given to the problem where a decision must be made about the quantity of a resource to purchase for a single
location in the face of a fixed cost $C$ for each unit of the resource and a loss $L$ that will be incurred for each
unit of unmet need. This leads to the quantile score for the probability level $\tau = 1 - C/L$. From this point, the
WIS or CRPS can be obtained by averaging across a range of decision making settings with different cost and loss
parameters, using a similar motivation that we used to obtain the IAS from the AS in section
\ref{sec:methods.detailed.integrated_allocation} \citep{gneiting2011weightedScoringRules}.

\section{Evaluating forecasts of COVID hospitalizations using the allocation score}
\label{sec:application}

We illustrate with an application to hospital admissions in the U.S., considering a hypothetical problem of allocating a
limited supply of medical resources to states.

\subsection{Data}

%\subsubsection{Hospitalization data}
%Starting in the summer of 2020, the US Health and Human Services (HHS) began reporting counts of daily new admissions to hospitals for individuals with COVID-19 \citep{healthdatagov_covid-19_nodate}.
%These daily counts were available for the US as a whole, and all states and several additional jurisdictions such as Puerto Rico and Washington DC.
%The data were updated daily and were available for download by the public through the HHS HealthData.gov website.
%For this analysis, we downloaded the hospitalizations data through the covidHubUtils R package, which connects users to the most recent version of the data \cite{wang-covidhubutils}.

%\subsubsection{Forecast data}
The US COVID-19 Forecast Hub collected short-term forecasts of daily new hospital admissions for individuals with COVID-19 starting in December 2020 \citep{cramer_united_2022}.
The target data for these forecasts were hospital admissions as reported by the US Department of Health and Human Services through the HealthData.gov website.
Forecasts were probabilistic predictions of the number of new hospital admissions on a particular day in the future, in a specific jurisdiction of the US (national level, state, or territory).
Probability distributions were represented using a set of 23 quantiles for each individual prediction, including a median and the lower and upper limits of 11 central prediction intervals, from a 99\% to a 10\% prediction interval.

The analysis in this work focuses on forecasts made before and during the first wave of the Omicron SARS-CoV-2 variant in the US.
As such, we analyzed forecasts for the 15 weeks starting with Monday November 22, 2021 through Monday February 28, 2022.
%Forecast data were downloaded on for analysis on format(as.Date(metadata$time[which(metadata$name == "forecast_data")]), "%B %d, %Y").

Submission to the Forecast Hub followed a weekly cycle, and each Monday the Hub collected the most recent forecasts submitted by all teams that met certain inclusion criteria and created ensemble forecasts using quantile averaging \citep{ray_comparing_2023}.
Our analysis includes these ensembles (COVIDhub-ensemble and COVIDhub-trained\_ensemble) as well as one other ensemble of hub models created by another team (JHUAPL-SLPHospEns) and several other individual models.
Models were eligible to be included in the analysis if they were designated as a ``primary'' model from a team.
For a model to have a complete, eligible submission in a given week, it had to have a 14 day-ahead forecast for all 50 states plus Washington DC.
Models had to have a complete forecast for at least 4 of the 15 weeks in the analysis to be eligible for inclusion.

The hospitalization data used for scoring forecasts were downloaded on February 07, 2024.

\subsection{Evaluation metrics}

We measured forecast skill using two forecast scores, the allocation score (AS) and the weighted interval score (WIS), both defined above.
We computed these scores for the 14 day ahead forecasts made each week.
For this analysis, we fixed the resource constraint used for the AS to be $K=15,000$, based roughly on a reported number of ventilators available for reallocation in the US \citep{ajao_assessing_2015}.
We computed the mean WIS (MWIS) across all of the forecasted state-level locations.
%For each week, we computed the allocation score for the 14 day-ahead forecast.

For both scores, we also computed models standardized rank among all models that submitted forecasts each week. The standardized rank is between 0 and 1, where 0 corresponds to the worst rank and 1 to the best. In the case of a tie between one or more models, all models received the better rank.

As described above, predictions were submitted to the Forecast Hub in the form of a set of 23 quantiles of the predictive distribution. The WIS can be directly calculated from these quantiles. However, calculation of the AS requires that the full cumulative distribution functions of the forecast distributions for each location are available. For the purpose of this analysis, to calculate the allocation score we approximated the full cumulative distribution functions based on the provided quantiles. On the interior of the provided quantiles we used a monotonic cubic spline to interpolate the quantiles, and in the lower and upper tails we used normal distributions with parameters selected so as to match the two lowest and two highest quantiles (see section \ref{sec:distfromq} the supplement for further details). In 
section \ref{sec:propriety_of_parametric_approximation} of the supplement, we show that if this evaluation procedure had been specified prospectively, the resulting score would be proper \textemdash but a post hoc application of this procedure is improper. We use the procedure here to illustrate the properties of the AS, and note that a collaborative forecasting hub interested in using the AS for evaluation could circumvent issues with propriety by communicating the definition of the AS to forecasters and then collecting allocations at specified resource levels as part of forecast submissions.

%We also computed a standardized rank for the allocation score for each model $m$ and week $w$.
%First, we computed the number of models that forecasted that week ($n_w$) and the rank of model $m$ among the $n_w$ models ($r_{m,w}^{AS}$).
%The model with the best allocation score received a rank of 1 and the worst received a rank of $n_w$.
%In the case of a tie between one or more models, all models received the better rank.
%We then rescaled these rankings to compute the allocation score standardized rank ($sr_{m,w}^AS$) between 0 and 1, where 0 corresponds to the worst rank and 1 to the best.

%\begin{equation}
%sr_{m,w}^{AS} = 1 - \frac{r^{AS}_{m,w}-1}{n_w-1}
%\end{equation}

%\subsubsection{Weighted Interval Score (WIS)}

%The Weighted Interval Score (WIS), as described in earlier sections, measures the alignment of a single probabilistic forecast ($F$) with an observation ($y$).
%We computed the mean WIS across all $L$ locations for each model and each forecasted week as
%\begin{equation}
%MWIS_{m,w} = \frac{1}{L}\sum_{l=1}^L WIS(F_{l,m,w},y)
%\end{equation}
%where $F_{l,m,w}$ is the probabilistic forecast from model $m$ for location $l$ and week $w$.
%Using the same procedure as for allocation scores described above, we computed standardized ranks for MWIS ($sr_{m,w}^{MWIS}$).

\subsection{Allocation benchmarks}

We used two benchmark methods as references to allow for comparison of model-derived allocations with existing standard practices.
First, we evaluated forecasts from the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk \citep{cramer_evaluation_2022}.
Second, we generated a proposed set of allocations where the quantity allocated to each state was proportional to that state's population (using US Census data of vintage 2022 \citep{us_census_nst_est2022}), referred to below as \texttt{per-capita}.
These two approaches generally reflect common choices for ``best practices'' of allocation: either allocating resources based on the most recent observed data, or in proportion to the population of each location.
%As such, they serve as good comparisons for allocations based on forecast models.

\subsection{Data and code availability}

All forecast data used in this evaluation are available through the COVID-19 Forecast Hub \citep{cramer_united_2022}. An R package implementing the allocation score is available at \url{https://github.com/aaronger/alloscore}. All code and data for the analyses presented in this manuscript are available at \url{https://github.com/aaronger/utility-eval-papers}. The analyses were generated using a reproducible workflow using R version 4.3.1 (2023-06-16) and the {\tt targets} package \citep{Rcore-2023, landau_2021_targets}.

\subsection{Application results}

\subsubsection{Anatomy of forecast scores for one week}

To illustrate the mechanics of allocation scoring, we start by focusing on how forecasts generated on or before December 20, 2021, with predictions for January 03, 2022, were scored by different metrics.
This week was around the peak of the Omicon wave nationally, with individual states typically observing a peak at or after January 3, 2022.

%For many locations, forecasts predicted lower values than were eventually observed, as this was during the period of steep increase of viral transmission across many states.
Of the 10 models evaluated for this one week, the CU-select model had the most accurate forecasts according to the allocation score while the USC-SI\_kJalpha model had the most accurate forecasts based on MWIS (Table \ref{tab:multi-k-scores}).
The JHUAPL-Bucky model had the second best MWIS but the third worst allocation score.

% latex table generated in R 4.3.1 by xtable 1.8-4 package
% Wed Feb 14 01:08:30 2024
\begin{table}[ht]
\centering
\begin{tabular}{lrrrr}
  \hline
Model & AS & MWIS & IAS centered at 15k & IAS uniform \\ 
  \hline
CU-select & 669 & 133 & 774 & 326 \\ 
  per-capita & 865 & - & 1029 & 366 \\ 
  COVIDhub-ensemble & 873 & 159 & 1067 & 438 \\ 
  USC-SI\_kJalpha & 995 & 91 & 1216 & 1097 \\ 
  JHUAPL-Gecko & 1034 & 164 & 1141 & 418 \\ 
  MUNI-ARIMA & 1084 & 169 & 1248 & 440 \\ 
  COVIDhub-trained\_ensemble & 1089 & 169 & 1271 & 823 \\ 
  COVIDhub-baseline & 1175 & 170 & 1317 & 535 \\ 
  JHUAPL-Bucky & 1358 & 102 & 1566 & 1214 \\ 
  JHUAPL-SLPHospEns & 1540 & 129 & 1604 & 1102 \\ 
  UVA-Ensemble & 2469 & 213 & 2635 & 2494 \\ 
   \hline
\end{tabular}
\caption{For one illustrative week, a comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS). All metrics are shown for 10 models that made forecasts of hospital admissions for 2022-01-03. Results are sorted by AS. For all metrics, lower scores indicate better accuracy.} 
\label{tab:multi-k-scores}
\end{table}

%Forecasts and allocations for a selection of states with high numbers of hospitalizations on January 3, 2022 reveal mechanics about how allocations are made (Figure \ref{fig:thermometer-plot}).
%As described in the methods above, allocations for a given location from a specific model are generated by finding the single quantile of the probabilistic forecast across all locations that reaches the resource allocation limit (in this case, assumed to be 15,000).
%A direct comparison between the forecasts from the JHUAPL-Bucky and CU-select models show that while the JHUAPL-Bucky forecast distributions were closer to the eventual observations in many states, the allocations suggested by those forecasts often were more inefficient than those from the CU-select model (Figure \ref{fig:thermometer-plot}).

%For JHUAPL-Bucky, the allocations were assigned at the Buck_quantileth percentile of the predictive distribution and for CU-select they were at the CU_quantileth percentile.
%This reflects that in general the JHUAPL-Bucky forecasts assigned more probability to higher values and thus a low quantile value across all locations reached the allocation limit of 15,000.
%In California, New York and Texas, the eventually observed number of hospital admissions was closer to the predicted median from JHUAPL-Bucky than the median from CU-select, and was contained in JHUAPL-Bucky's 80\% PI but not in the 80\% PI from CU-select. However for each of these three states, the resource allocation was lower for JHUAPL-Bucky than for CU-select and therefore resulted in more hypothetical unmet need.
%In Florida, which was the state that saw the largest number of reported hospitalizations on this target date, the forecast from JHUAPL-Bucky missed by a wide margin (as did the forecast from CU-select) and the allocation was FL_diff units lower than the allocation derived from the CU-select model.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/thermometer-plot-1} \caption[Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the ten highest hospitalization counts]{Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the ten highest hospitalization counts. For each state, the dark black line shows the data observed when the forecast was made and the grey line shows eventually observed counts. The side-by-side shaded regions show the median (solid red horizontal line) and 50\%, 80\% and 95\% prediction intervals for the two selected models. The forecasts were made for new hospitalizations on January 3, 2022 (vertical dashed line, with number of hospitalizations indicated by red horizontal line segment at the intersection of the dashed line and the grey line of data). The vertical bars with purple, blue and red shading show the allocations. The purple bar goes from zero to the amount of need that was met by the allocation from that model to a specific location. A red bar indicates need that exceeded the resource allocation for a location, and a light blue bar shows the amount by which the resource allocation suggested by that model exceeded the need for that location.}\label{fig:thermometer-plot}
\end{figure}

\end{knitrout}

A comparison of the forecasts from the JHUAPL-Bucky and CU-select models shows that while the JHUAPL-Bucky forecast distributions were closer to the eventual observations in many states, the allocations suggested by those forecasts often were more inefficient than those from the CU-select model (Figure \ref{fig:thermometer-plot}).
In the example of this one week, JHUAPL-Bucky had a worse allocation score than CU-select because its forecasts led to allocations that sent excess resources to several states, such as Ohio, Pennsylvania, and Michigan, that would have been more effectively allocated to states that did receive enough resources to meet their needs, such as Florida and California (Figure \ref{fig:multi-loc-ranks}A). These allocation errors resulted because that model’s forecasts did not consistently capture the relative resource needs across different states. (Figure \ref{fig:thermometer-plot} \& \ref{fig:multi-loc-ranks}A). The CU-select model made some similar errors — most prominently, over-allocating resources to Ohio — but overall, it did a better job of forecasting the relative resource demands across different locations.

On the other hand, CU-select had worse performance as measured by WIS. Its forecasts were biased downwards, and it consistently incurred a large penalty for underprediction (Figure \ref{fig:multi-loc-ranks}B). Predictions from the JHUAPL-Bucky model were wider, and included the observed level of daily hospital admissions more often. Therefore, it did not receive as severe penalties for forecasts that underpredicted or overpredicted the actual hospitalization count.

%Together, these results corroborate the understanding that the allocation score and weighted interval score measure different aspects of forecast performance. The allocation score penalizes models that mischaracterize the relative magnitude of resource need across different locations, while WIS penalizes forecasts that do not capture the absolute magnitude of the target quantity in each location.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/multi-loc-ranks-1} \caption[Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (JHUAPL-Bucky and CU-select)]{Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (JHUAPL-Bucky and CU-select). Panel A shows the observed resource need, in this case the observed number of hospitalizations, for each state, along with the hypothetical number of resources allocated to the given location based on the forecasts from each model. The number of available resources was fixed at 15,000 and forecasts from each model were used to determine an optimal allocation strategy before the resource need was known. For most locations the resource need exceeded the resources allocated, indicated by the `observed resource need' bar being larger than the `allocation' bar. Panel B shows the breakdown of the weighted interval score (WIS) into components of underprediction, overprediction and dispersion. Larger values of WIS indicate more error, and the full WIS score for each location can be decomposed into the three components shown here.}\label{fig:multi-loc-ranks}
\end{figure}

\end{knitrout}

\subsubsection{Forecast scores showed differences in aggregate and over time}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/metrics-over-time-1} \caption[Hospital admissions and evaluation metrics over time]{Hospital admissions and evaluation metrics over time. Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022. These are the dates for which forecasts were made and evaluated. A horizontal dashed line at 15,000 shows the hypothetical resource constraint $K$. Panel B shows allocation scores (AS) for each model's 14 day-ahead forecast, across all US states. The $x$-axis corresponds to the date of the observation that a model's prediction was targeting (e.g., the date the forecast was made plus the forecast horizon). AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down). In Panel B, the per-capita allocation is drawn in a heavier solid line. Panel C shows the MWIS metric across weeks, averaged across all states. MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the peak of the Omicron wave.}\label{fig:metrics-over-time}
\end{figure}

\end{knitrout}

Allocation scores varied substantially by date and by model (Figure \ref{fig:metrics-over-time}).
For predictions made for the first three Mondays in December 2021 and the last three Mondays in February 2022 all models had allocation scores under 500 (and the mean across all models was less than 100), indicating that unnecessary unmet need was fairly low on those days.
The allocation scores were on the whole highest when the observed number of new hospital admissions was closest to the resource threshold of 15,000, as those are the times when any mistakes in allocation are costly in terms of wasting resources in one location that could have been used in another.
%Predictions made during the peak week and just after showed the highest variation in allocation scores, with some models having allocation scores under 1000 and others having values over 3500.

%Overall, across the first 11 weeks evaluated (we excluded the last four since nearly all the models achieved an allocation score of zero), the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk, had the highest rank for allocation score in four weeks, more than any other model except the \texttt{COVIDhub-ensemble} which also had four.

%Mean weighted interval scores (MWIS) also varied by date and model, and more clearly were dependent on the scale of the observed data.
%MWIS values were low (all models under 100) for all Mondays in December 2021 and the final four Mondays evaluated.
%Across all models both the average and median MWIS value for every Monday in January was above 100, with the largest errors occurring one and two weeks after the peak was observed.

Averaging across all weeks, the ensemble forecast achieved a better mean allocation score (MAS) than two benchmark methods.
The COVIDhub-ensemble had the best MAS across all weeks, the per-capita allocation had the second best MAS, and the COVIDhub-baseline model had the sixth best allocation (Table \ref{tab:multi-week-performance-summary}).
No individual model had a better MAS than the per-capita allocation.

Most models were ranked similarly when evaluated using mean allocation scores (MAS) and mean weighted interval scores (MWIS), with some places of disagreement.
The the four most accurate and the five least accurate models were the same according to both metrics, although not in exactly the same order (Table \ref{tab:multi-week-performance-summary}).
(This comparison excludes the per-capita method which can be used to suggest allocations but does not make probabilistic forecasts and therefore cannot have a WIS.)
Notably, the best model according to MWIS was the JHUAPL-SLPHospEns model, but this model only had the fourth best allocation scores.

% latex table generated in R 4.3.1 by xtable 1.8-4 package
% Wed Feb 14 01:08:35 2024
\begin{table}[ht]
\centering
\begin{tabular}{lrr}
  \hline
model & MAS & MWIS \\ 
  \hline
COVIDhub-ensemble & 389 & 70 \\ 
  per-capita & 464 & - \\ 
  COVIDhub-trained\_ensemble & 483 & 87 \\ 
  CU-select & 502 & 81 \\ 
  JHUAPL-SLPHospEns & 526 & 67 \\ 
  COVIDhub-baseline & 594 & 93 \\ 
  JHUAPL-Bucky & 643 & 112 \\ 
  MUNI-ARIMA & 707 & 116 \\ 
  JHUAPL-Gecko & 929 & 110 \\ 
  USC-SI\_kJalpha & 1473 & 155 \\ 
   \hline
\end{tabular}
\caption{Mean weighted interval scores (MWIS) and mean allocation scores (MAS) by model across 13 weeks. For the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, plus a per-capita allocation, these results show the average performance of each model across all weeks. Lower scores indicate better performance, and models are sorted with the most accurate allocations at the top. WIS is not defined for the per-capita model as this approach only provides a mechanism to allocate resources, it does not provide probabilistic forecasts.} 
\label{tab:multi-week-performance-summary}
\end{table}

\subsubsection{Metrics were not consistently correlated over time}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/metrics-correlation-1} \caption[Association of standardized ranks for MWIS and allocation score by model and week]{Association of standardized ranks for MWIS and allocation score by model and week. Each facet of the plot corresponds to one model. Within each facet, each point corresponds to a week. The $x$- and $y$-values correspond to the MWIS standardized rank and the allocation score standardized rank for that week. Points corresponding to earlier dates have darker shading. The size of the point corresponds to the observed value on the date for which the prediction was made. Models show different degrees of association between the two metrics.}\label{fig:metrics-correlation}
\end{figure}

\end{knitrout}

Models showed differing levels of correlation between their allocation scores and MWIS values (Figure \ref{fig:metrics-correlation}), explaining in part why aggregate MWIS and MAS values did not always agree.
Here are some examples of the different model-specific patterns observed:
\begin{itemize}
\item Several models showed a positive association between allocation score and MWIS ranks (e.g., \texttt{Karlen-pypm} and \texttt{USC-SI\_kJalpha}).
\item One model had consistently strong MWIS ranks but also had highly variable allocation score ranks with no clear association between the two (e.g., \texttt{JHUAPL-SLPHospEns})
\item One model performed consistently well for both metrics (\texttt{COVIDhub-ensemble})
\item One model performed consistently well for allocation score but had only middling ranks for MWIS (\texttt{CMU-TimeSeries})
\end{itemize}

\subsubsection{Integrated allocation score across values of K}

%The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the constraint ($K$), possibly taking weights into account for different values of $K$ that might be more or less likely (Section \ref{sec:methods.detailed.integrated_allocation}).
%This could be useful in situations where the actual constraint may not be known precisely, or where we wish to consider results across a range of decision making settings with different constraints.
%Unlike in the above analyses where we conducted analyses assuming that $K$ was known to be 15,000, the analyses in this subsection are conducted assuming different distributions on $K$.

AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of hospitalizations on January 3, 2021 as well as the per-capita allocation (Figure \ref{fig:multi-k}A).
These calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of 19,581 that day.
Rankings of AS from all models were fairly stable across a range of values for K, with some crossings, especially at $K$ values further away from the observed value.

The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the constraint ($K$), possibly taking weights into account for different values of $K$ that might be more or less likely (Section \ref{sec:methods.detailed.integrated_allocation}).
IAS was computed for two distributions on $K$, one uniform across the entire range and the other a symmetric distribution around 15,000 (Figure \ref{fig:multi-k}B).
Both versions of the IAS were correlated with the original AS conducted at $K=15,000$, with the higher correlation coming from the distribution that was centered at $K=15,000$ (Figure \ref{fig:multi-k}C).
Model rankings based on the AC and the centered IAS were roughly similar, with the top and bottom three approaches being the same for both scores (Table \ref{tab:multi-k-scores}).

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/multi-k-1} \caption[Allocation scores (AS) across different resource constraints ($K$) for 10 models that made forecasts on 2021-12-20]{Allocation scores (AS) across different resource constraints ($K$) for 10 models that made forecasts on 2021-12-20. Panel A shows, for each model, the AS for values of $K$ between 200 and 60,000 at increments of 200. The AS show a sharp peak just under 20,000, near the eventually observed number of hospitalizations. Panel B shows two possible weighting functions for the Integrated Allocation Score (IAS). The first (dark blue circles) computes weights proportional to a normal distribution centered at 15,000 (solid vertical gray line) with a standard deviation of 3,000, and truncated to be between 5,000 and 25,000. The second (light orange triangles) uses a uniform weight for all possible values of $K$. Note that the AS used in earlier sections of the application uses the single fixed value of $K=15,000$.  Panel C shows how either method of computing the IAS ($y$-axis) is correlated with AS at $K=15,000$ ($x$-axis). Every point represents the AS and IAS for one model. The IAS centered at 15,000 are more closely correlated with the AS values.}\label{fig:multi-k}
\end{figure}

\end{knitrout}

\section{Discussion}
\label{sec:discussion}

In epidemiological forecasting, well-known proper scoring rules such as the log score or variations on the continuous ranked probability score (such as the weighted interval score, or WIS) have been frequently utilized to evaluate probabilistic forecasts.
Often models are ranked with respect to accuracy according to a particular score, but without reference to any underlying decision-making process for which that score was designed or from which it might be derivable as a Bayes scoring rule.
With careful thought and collaboration between modelers and public health officials, we argue that scores that are more aligned with public health decisions could be developed to inform specific problems.
We have demonstrated that forecast evaluation methods that are tied to a specific decision making context can yield model rankings that differ substantively from those based on standard measures of forecast skill, a result that aligns with findings in other fields \citep{leitch1991economicForecastEval, murphy1993whatisagoodforecast, cenesizoglu2012returnPredictionEconValue}.
Additionally, we show that an existing ensemble forecast approach was the only method to outperform two benchmark allocation approaches in a hypothetical application.

%In particular, the allocation score penalizes models that mischaracterize the relative magnitude of resource need across different locations, while WIS penalizes forecasts that do not capture the absolute magnitude of the target quantity in each location.

%We often conceive of infectious disease forecasts as being useful for decision making purposes, but it is rare for forecast evaluation to be tied directly to the value of the forecasts for informing those decisions. This work seeks to address that gap.
%However, we do note that the decision-making context presented in this work, while motivated by examples from real-world public health resource allocation problems, has not been used to inform an actual real-time decision-making process.

While our example used a hypothetical scenario to illustrate a potential application of the allocation score, resource allocation decisions are a realistic example of the kinds of decisions that could motivate more targeted forecasting exercises.
%One real-world example, which was the motivating example presented in the work above, is allocation of ventilators during a respiratory viral pandemic\citep{huang_stockpiling_2017}.
We used ventilators as a notional example for a resource allocation problem because it (a) has been identified as a scarce resource during outbreaks \citep{huang_stockpiling_2017}, (b) could be allocated across different locations, with logistical considerations depending on the geographical scales, and (c) would likely be correlated (although not perfectly so) with hospitalization data.
However, this example has limitations in that not everyone who is hospitalized needs a ventilator, and new hospital admissions may not be the best stand in for a need that corresponds to a prevalence of severe cases.
Other possible real-world examples include the allocation of a limited stockpile of vaccinations\citep{araz_geographic_2012,persad_fair_2023} or diagnostic tests\citep{du_optimal_2022,pasco_covid-19_2023}.
While there may be limitations in this specific pairing of a resource and forecasted outcome, we emphasize that the important result from this work is how evaluating forecasts using a fundamentally different kind of score can illuminate previously hidden aspects of value in forecasts.
In this setting, the feature of a forecast that emerges as being important is that the probabilistic forecasts predict the relative magnitudes between locations accurately, even if the absolute magnitudes are inaccurate.

In practice, there are many potential users of forecasts with many different decision making problems.
While outcomes of some decisions could potentially be specified by loss functions, other uses for forecasts may be less quantifiable, such as when forecasts are used to inform a general sense of situational awareness that may feed in unmeasurable ways into other decisions.
Even those that are easily framed as expected loss minimization may differ enough that no single score would be
appropriate for all users.
Ideally, targeted forecasting tools could be developed through close collaboration between modelers and public health officials.
However, this may only be possible in settings with sufficient staffing on both an analytics and a public health team.
Increasingly, collaborative modeling hubs are being used to generate ``one-size-fits-all'' forecasts for many locations at once.
In these settings, where tailored models are not available, it still could be possible to evaluate contributed models using a set of multiple scores to support public health end users in understanding the value of forecasts as an input to their particular decision making contexts.
% NGR: leaving this out for now as it feels like too much detail?
% This may be tricky to operationalize in the setting of a general forecast hub. It matters how you elicit and represent probabilistic forecasts (quantiles? samples? cdfs?).

%In our specific application, we made several key observations that we believe should inform future work in this area.
%First, it is clear that different metrics (in our case allocation score and mean weighted interval score) captured different aspects of forecast performance.
%The allocation score penalizes models that mischaracterize the relative magnitude of resource need across different locations, while WIS penalizes forecasts that do not capture the absolute magnitude of the target quantity in each location.
%A forecast with a good allocation score can have a poor weighted interval score and vice versa.
%The mean WIS (MWIS) was strongly dependent on the scale of the forecasted quantity (e.g., when hospitalizations were high, so were MWIS scores).
%The allocation score was not as scale-dependent, as the highest/worst allocation scores were observed when the number of hospitalizations was closest to the allocation constraint.
%We also note that contributed models found it hard to achieve a better allocation score than a na\"ive baseline model that just predicted a flat line from the last previous observation with wide uncertainty.
%This suggests suggests that contributed models (other than the ensemble, which combined forecasts from all contributed models) were not consistently adding value to allocation decisions over just extrapolating from the current levels.

There are several important limitations to the current work.
%NGR - removing the sentence below because it seems like not a central concern/limitation, and I was triyng to cut. Happy to be overruled.
%The allocation score we developed here does not directly account for important considerations such as fairness or equity of allocations.
The proposed framework does not attempt to capture the broader context of decision making. For example, in practice it may be possible to increase the resource constraint $K$ by shifting funding from other disease mitigation measures.
We also note that in some settings, a ``successful'' epidemiological forecast may lead to policy decisions that change the distribution of the predicted outcome $Y$. Our framework would need considerable enhancements before being applicable to forecast evaluation beyond horizons
for which causal feedback can be neglected.

An opportunity for further investigation is to more carefully evaluate whether a forecast adds value to existing decision making processes.
In the context of decisions about allocations, standard procedures might involve extrapolating need based on a current observed data (similar to the benchmark approaches presented above), with or without adjustments based on other political or real-world considerations.
For example, in many settings public health stakeholders will make decisions after synthesizing information from a variety of quantitative and qualitative sources coupled with expert judgment.
The allocation score presented in this work does not directly measure whether a given forecast adds useful information to such an existing decision-making process.
While the scoring procedures as presented do not directly address this question, they could be modified (say, by comparison to a `baseline' model or expert-elicited allocations in the absence of forecast data) to quantify the benefit of using a forecast to inform a specific decision.

%We see this work as an initial overture for what we hope will grow to be a large, collaborative body of work more closely coupling applied epidemiological forecasting with public health decision making.
%We note that a few papers have begun exploring similar linkages as described in the literature review{\textendash}but we see much room for additional work in this area.
%In some situations, individual or ensemble models could be developed to optimize scores that are attuned to a particular decision making setting.
%This area is also largely unexplored in the realm of public health to date, with some initial methodological development in econometrics \citep{loaiza-maya_focused_2021}.

In conclusion, we argue that the way modelers and policymakers view and evaluate forecasts should change depending on the specific decision-making context.
Defaulting to standard forecast evaluation metrics can mask the utility (or disutility) of certain forecasts, or lead to forecasts being used in decision making contexts very different from those for which they might offer useful guidance.
New collaborative work between public health officials and modeling teams is needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or simulation exercises that could be used to inform further development of new or alternative scoring metrics.
We see this work as an initial overture for what we hope will grow to be a large, collaborative body of work more closely coupling applied epidemiological forecasting with public health decision making.

\section*{Acknowledgements}

We wish to thank the following individuals who contributed valuable comments and feedback on early versions of this work: Matthew Biggerstaff, Rebecca Borchering, Sebastian Funk, Melissa Kerr, and Jeffrey Shaman.

This work has been supported by the National Institutes of General Medical Sciences (R35GM119582) and the U.S. CDC(1U01IP001122). The content is solely the responsibility of the authors and does not necessarily represent the official views of NIGMS, the National Institutes of Health, or CDC.

\bibliography{allocation}

\end{document}